---
title: "Keratinocytes removal - test dissociation 3h vs 18h"
subtitle: "Keratinocytes removal"
author: "Pernelle"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

```{r clean, echo = FALSE}
rm(list = ls())
```

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <= Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


This file is used to generate a dataset containing only fibroblasts from different mouse skin datasets in cNFs project. It investigates sub-populations within this population, and perform differential expression analysis.

```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
```


# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "woKR"
out_dir = "."
```

These are samples of interest :

```{r project_names_oi}
project_names_oi = c("2023_10","2023_09","2023_12","2023_11")
```


We load the sample information :

```{r custom_palette_sample, fig.width = 10, fig.height = 1}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "dissociation_methods_metadata_sample_info.rds"))

if (project_names_oi == "all") {
  project_names_oi = sample_info$project_name
} else {
  sample_info = sample_info %>%
    dplyr::filter(project_name %in% project_names_oi)
}

ggplot2::ggplot(sample_info) +
  ggplot2::geom_label(aes(x = project_name, y = 0, label = project_name, fill = project_name)) +
  ggplot2::geom_text(aes(x = project_name, y = 1, label = sample_identifier)) +
  ggplot2::lims(y = c(-1, 2)) +
  ggplot2::scale_fill_manual(values = sample_info$color, breaks = sample_info$project_name) +
  ggplot2::theme_void() +
  ggplot2::theme(legend.position = "none")
```

Here are custom colors for each cell type :

```{r color_markers, fig.width = 10, fig.height = 1, class.source = "fold-hide"}
color_markers = readRDS(paste0(out_dir, "/../1_metadata/", "dissociation_methods_metadata_color_markers.rds"))
cell_markers = readRDS(paste0(out_dir, "/../1_metadata/", "dissociation_methods_metadata_cell_markers.rds"))

data.frame(cell_type = names(color_markers),
           color = unlist(color_markers)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank())
```


# Make `r save_name` dataset

## Individual datasets

We extract a cluster annotated as `r save_name` in each dataset. For each sample, we :

* load individual dataset
* look at `r save_name` markers
* look at `r save_name` score for each cluster
* select the best clusters based on average `r save_name` score

We load individual datasets :

```{r sobj_list}
sobj_list = lapply(project_names_oi, FUN = function(project_name) {
  subsobj = readRDS(paste0(out_dir, "/../2_individual/datasets/", project_name, "_sobj_filtered.rds"))
  return(subsobj)
})
names(sobj_list) = project_names_oi

lapply(sobj_list, FUN = dim)
```
We change the resolution of each dataset :

```{r resolution}
sobj_list = lapply(sobj_list, FUN = function(one_sobj) {
  one_sobj = Seurat::FindClusters(one_sobj, resolution = 0.3)
  return(one_sobj)
})
```

We represent cells in the tSNE :

```{r name2D}
name2D = "RNA_pca_20_umap"
```


```{r cluster_visualisation, fig.width = 12, fig.height = 13}
plot_list = lapply(sobj_list, FUN = function(one_sobj) {
  subtitle = as.character(unique(one_sobj$project_name))
  p = Seurat::DimPlot(one_sobj, reduction = name2D,
                      group.by ="seurat_clusters", label = TRUE) +
    ggplot2::labs(subtitle = subtitle) +
    Seurat::NoAxes() +
    Seurat::NoLegend() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.subtitle = element_text(hjust = 0.5))
  return(p)
})

patchwork::wrap_plots(plot_list, ncol=4)



plot_list = lapply(sobj_list, FUN = function(one_sobj) {
  subtitle = as.character(unique(one_sobj$project_name))
  p = Seurat::FeaturePlot(one_sobj, reduction = name2D,
                      features = c("Krt14")) +
    ggplot2::labs(subtitle = subtitle) +
    Seurat::NoAxes() +
    Seurat::NoLegend() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.subtitle = element_text(hjust = 0.5))
  return(p)
})

patchwork::wrap_plots(plot_list, ncol=4)

plot_list = lapply(sobj_list, FUN = function(one_sobj) {
  subtitle = as.character(unique(one_sobj$project_name))
  p = Seurat::FeaturePlot(one_sobj, reduction = name2D,
                      features = c("Krt1")) +
    ggplot2::labs(subtitle = subtitle) +
    Seurat::NoAxes() +
    Seurat::NoLegend() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.subtitle = element_text(hjust = 0.5))
  return(p)
})

patchwork::wrap_plots(plot_list, ncol=4)

plot_list = lapply(sobj_list, FUN = function(one_sobj) {
  subtitle = as.character(unique(one_sobj$project_name))
  p = Seurat::FeaturePlot(one_sobj, reduction = name2D,
                      features = c("Epcam")) +
    ggplot2::labs(subtitle = subtitle) +
    Seurat::NoAxes() +
    Seurat::NoLegend() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.subtitle = element_text(hjust = 0.5))
  return(p)
})

patchwork::wrap_plots(plot_list, ncol=4)


plot_list = lapply(sobj_list, FUN = function(one_sobj) {
  subtitle = as.character(unique(one_sobj$project_name))
  p = Seurat::FeaturePlot(one_sobj, reduction = name2D,
                      features = c("Krt15")) +
    ggplot2::labs(subtitle = subtitle) +
    Seurat::NoAxes() +
    Seurat::NoLegend() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.subtitle = element_text(hjust = 0.5))
  return(p)
})

patchwork::wrap_plots(plot_list, ncol=4)
```

We add a new column to then select cells.

We attribute a score for different cell types in each dataset :


We represent, on the projection, score, clusters we will select, and represent the average score for each cluster, by dataset.


```{r score_keratinocytes}

## Keratinocytes markers

kerato_markers = c("Krt28","Cdh1","Epcam","Krt15")

## Keratinocytes score attribution

sobj_list = lapply(sobj_list, FUN = function(one_sobj) {
  one_sobj$kerato_score = Seurat::AddModuleScore(
    one_sobj,
    features = list(kerato_markers))$Cluster1
  
  return(one_sobj)
})

## Keratinocytes selection

mean_score_thresh_kerato = 0.15

sobj_list = lapply(sobj_list, FUN = function(one_sobj) {
  ## Score by cluster
  score_by_clusters = one_sobj@meta.data %>%
    dplyr::select(seurat_clusters, kerato_score) %>%
    dplyr::group_by(seurat_clusters) %>%
    dplyr::summarise(avg_score = mean(kerato_score)) %>%
    as.data.frame()
  
  ## Cluster of interest
  cluster_of_interest = score_by_clusters %>%
    dplyr::filter(avg_score > mean_score_thresh_kerato) %>%
    dplyr::pull(seurat_clusters) %>%
    as.character()
  one_sobj$isnot_kerato = ifelse(one_sobj$seurat_clusters %in% cluster_of_interest,
                                   yes = FALSE, no = TRUE)
  
  ## Output
  return(one_sobj)
})

lapply(sobj_list, FUN = function(one_sobj) {
  table(one_sobj$isnot_kerato)
}) %>% do.call(rbind, .)

```

```{r score_remove_kerato, fig.width = 12, fig.height = 30, class.source = "fold-hide"}

plot_list = lapply(sobj_list, FUN = function(one_sobj) {
  project_name = as.character(unique(one_sobj$project_name))
  
  #--------------------- copied-pasted
  ## Score by cluster
  score_by_clusters = one_sobj@meta.data %>%
    dplyr::select(seurat_clusters, kerato_score) %>%
    dplyr::group_by(seurat_clusters) %>%
    dplyr::summarise(avg_score = mean(kerato_score)) %>%
    as.data.frame()
  
  ## Cluster of interest
  cluster_of_interest = score_by_clusters %>%
    dplyr::filter(avg_score < mean_score_thresh_kerato) %>%
    dplyr::pull(seurat_clusters) %>%
    as.character()
  #---------------------
  
  ## Number of cells
  nb_cells_ind = sum(one_sobj$kerato_score < mean_score_thresh_kerato)
  nb_cells_clust = sum(one_sobj$seurat_clusters %in% cluster_of_interest)
  
  ## Figure
  p_score = Seurat::FeaturePlot(one_sobj, reduction = name2D,
                                features = "kerato_score") +
    ggplot2::labs(title = paste0(nb_cells_ind, " cells"),
                  subtitle = paste0("Score keratinocytes"),
                  y = project_name) +
    ggplot2::scale_color_gradient2(low = aquarius:::color_cnv[1],
                                   mid = aquarius:::color_cnv[2],
                                   high = aquarius:::color_cnv[3],
                                   midpoint = 0) +
    Seurat::NoAxes() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.subtitle = element_text(hjust = 0.5),
                   axis.title.y = element_text(size = 20, face = "bold", angle = 90))
  
  p_cluster = Seurat::DimPlot(one_sobj, reduction = name2D,
                              group.by = "isnot_kerato") +
    ggplot2::labs(title = paste0(nb_cells_clust, " cells"),
                  subtitle = paste0("Clusters : ", paste(cluster_of_interest, collapse = ", "))) +
    Seurat::NoAxes() + Seurat::NoLegend() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5))
  
  p_mean = ggplot2::ggplot(score_by_clusters, aes(x = seurat_clusters, y = avg_score)) +
    ggplot2::geom_point() +
    ggplot2::geom_hline(yintercept = mean_score_thresh_kerato, col = "red") +
    ggplot2::labs(x = "Cluster ID", y = "Mean score",
                  title = "Mean score by cluster") +
    ggplot2::theme_classic() +
    ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"))
  
  ## Output
  return(list(p_score, p_cluster, p_mean))
}) %>% unlist(., recursive = FALSE)

patchwork::wrap_plots(plot_list, ncol = 3, widths = c(1,1,2))
```




## Combined dataset

### cNFs project

For each sample, we extract cells of interest based on clusters mean score, and build a new dataset.

```{r subset_is_of_interest}


### Keratinocytes removal

sobj_list = lapply(sobj_list, FUN = function(one_sobj) {
  if (sum(one_sobj$isnot_kerato) > 0) {
    one_sobj = subset(one_sobj, isnot_kerato == TRUE)
  } else {
    one_sobj = NA
  }
  return(one_sobj)
})

lapply(sobj_list, FUN = dim)


```

We select project names for which cells are available :

```{r there_are_cells}
there_are_cells = lapply(sobj_list, FUN = ncol) %>%
  unlist() %>%
  names()
there_are_cells
```

We remove datasets without cells :

```{r subset_sobj_list}
sobj_list = sobj_list[there_are_cells]

lapply(sobj_list, FUN = dim)
```

We combine all datasets :

```{r merge_datasets}
sobj = base::merge(sobj_list[[1]],
                   y = sobj_list[c(2:length(sobj_list))],
                   add.cell.ids = there_are_cells)

rm(sobj_list)
sobj
```

We clean this dataset :

```{r clean_dataset}
column_of_interest = c("orig.ident", "nCount_RNA", "nFeature_RNA",
                       "log_nCount_RNA", "project_name", "mouse_id",
                       "age", "genotype",
                       "sample_identifier","condition",
                       "sample_type","percent.st","cell_type",
                       "Seurat.Phase","percent.mt","percent.rb",
                       "cyclone.Phase","dissociation")
sobj@meta.data[!(colnames(sobj@meta.data) %in% column_of_interest)] = NULL

head(sobj@meta.data)
```
We reset factor levels :

```{r reset_factor_levels}
#sample_info$dissociation = factor(sample_info$dissociation,
                             #     levels = sample_info$dissociation)
sample_info$project_name = factor(sample_info$project_name,
                                  levels = sample_info$project_name)
sample_info$sample_identifier = factor(sample_info$sample_identifier,
                                       levels = sample_info$sample_identifier)
sample_info$genotype = factor(sample_info$genotype,
                              levels = unique(sample_info$genotype))
sample_info$sample_type = factor(sample_info$sample_type,
                                 levels = unique(sample_info$sample_type))
sample_info$color = as.character(sample_info$color)

summary(sample_info)
```

This is metadata :

```{r see_meta.data}
head(sobj@meta.data)
```
We set factor levels

```{r sobj_set_factor_levels}
sobj$orig.ident = factor(sobj$orig.ident,
                         levels = levels(sample_info$project_name))
sobj$project_name = factor(sobj$project_name,
                           levels = levels(sample_info$project_name))
sobj$sample_identifier = factor(sobj$sample_identifier,
                                levels = levels(sample_info$sample_identifier))
sobj$genotype = factor(sobj$genotype,
                       levels = levels(sample_info$genotype))
sobj$sample_type = factor(sobj$sample_type,
                          levels = levels(sample_info$sample_type))

```

# Processing

We remove genes that are expressed in less than 5 cells :

```{r filter_genes}
sobj = aquarius::filter_features(sobj, min_cells = 5)
sobj
```

## Metadata

How many cells by sample ?

```{r table_orig_ident}
table(sobj$project_name)
```

We represent this information as a barplot :

```{r barplot_count, fig.width = 3, fig.height = 5}
aquarius::plot_barplot(df = table(sobj$project_name) %>%
                         as.data.frame.table() %>%
                         dplyr::mutate(x = "Fibroblasts"),
                       x = "x", y = "Freq", fill = "Var1",
                       position = position_fill()) +
  ggplot2::scale_fill_manual(values = sample_info$color,
                             breaks = sample_info$project_name,
                             name = "Sample of origin")
```
## Projection

We normalize the count matrix for remaining cells and select highly variable features in 10x Genomics data :

```{r normalization}
sobj = Seurat::NormalizeData(sobj,
                             normalization.method = "LogNormalize")

cells_fibro = colnames(sobj)
Seurat::VariableFeatures(sobj) = subset(sobj, cells = cells_fibro) %>%
  Seurat::FindVariableFeatures(., nfeatures = 2000) %>%
  Seurat::VariableFeatures()

sobj = Seurat::ScaleData(sobj)

sobj
```

We perform a PCA :

```{r pca}
sobj = Seurat::RunPCA(sobj,
                      assay = "RNA",
                      reduction.name = "RNA_pca",
                      npcs = 100,
                      seed.use = 1337L)
sobj
```

We choose the number of dimensions such that they summarize 60 % of the variability :

```{r ndims}
stdev = sobj@reductions[["RNA_pca"]]@stdev
stdev_prop = cumsum(stdev)/sum(stdev)
ndims = which(stdev_prop > 0.60)[1]
ndims
```

We can visualize this on the elbow plot :

```{r elbowplot, fig.width = 12, fig.height = 4}
elbow_p = Seurat::ElbowPlot(sobj, ndims = 100, reduction = "RNA_pca") +
  ggplot2::geom_point(x = ndims, y = stdev[ndims], col = "red")
x_text = ggplot_build(elbow_p)$layout$panel_params[[1]]$x$get_labels() %>% as.numeric()
elbow_p = elbow_p +
  ggplot2::scale_x_continuous(breaks = sort(c(x_text, ndims)), limits = c(0, 100))
x_color = ifelse(ggplot_build(elbow_p)$layout$panel_params[[1]]$x$get_labels() %>%
                   as.numeric() %>% round(., 2) == round(ndims, 2), "red", "black")
elbow_p = elbow_p +
  ggplot2::theme_classic() +
  ggplot2::theme(axis.text.x = element_text(color = x_color))

elbow_p
```

We generate a tSNE and a UMAP with `r ndims` principal components :

```{r tsne_umap}
sobj = Seurat::RunTSNE(sobj,
                       reduction = "RNA_pca",
                       dims = 1:ndims,
                       seed.use = 1337L,
                       reduction.name = paste0("RNA_pca_", ndims, "_tsne"))

sobj = Seurat::RunUMAP(sobj,
                       reduction = "RNA_pca",
                       dims = 1:ndims,
                       seed.use = 1337L,
                       reduction.name = paste0("RNA_pca_", ndims, "_umap"))
```

We can visualize the two representations. There is a batch effect :

```{r see_umap_tsne, fig.width = 8, fig.height = 4, class.source = "fold-hide"}
tsne = Seurat::DimPlot(sobj, group.by = "project_name",
                       reduction = paste0("RNA_pca_", ndims, "_tsne")) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name) +
  Seurat::NoAxes() + ggplot2::ggtitle("PCA - tSNE") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

umap = Seurat::DimPlot(sobj, group.by = "project_name",
                       reduction = paste0("RNA_pca_", ndims, "_umap")) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name) +
  Seurat::NoAxes() + ggplot2::ggtitle("PCA - UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

tsne | umap
```

There is a strong batch-effect. We remove batch-effect using Harmony :

```{r harmony, fig.width = 8, fig.height = 5}

library(Seurat)
set.seed(1337L)
my_harmony_embeddings <- harmony::HarmonyMatrix(
  data_mat  = as.matrix(sobj@reductions$RNA_pca@cell.embeddings),
  meta_data = sobj@meta.data,
  vars_use  = "dissociation",
  max.iter.harmony = 50,
  do_pca = FALSE
)
rownames(my_harmony_embeddings) <- rownames(sobj@reductions$RNA_pca@cell.embeddings)

#store the harmony reduction as a custom dimensional reduction called 'harmony' in the default assay
sobj[["harmony"]] <- Seurat::CreateDimReducObject(embeddings = my_harmony_embeddings, key = "harmony_", assay = DefaultAssay(sobj))

```

From this batch-effect removed projection, we generate a tSNE and a UMAP.

```{r harmony_tsne_umap, fig.width = 12, fig.height = 12}

sobj = Seurat::RunUMAP(sobj, 
                       seed.use = 1337L,
                       dims = 1:ndims,
                       reduction = "harmony",
                       reduction.name = paste0("harmony_", ndims, "_umap"),
                       reduction.key = paste0("harmony_", ndims, "umap_"))
sobj = Seurat::RunTSNE(sobj,
                       dims = 1:ndims,
                       seed.use = 1337L,
                       reduction = "harmony",
                       reduction.name = paste0("harmony_", ndims, "_tsne"),
                       reduction.key = paste0("harmony", ndims, "tsne_"))


```

We visualize the representation **after batch-effect correction** :

```{r see_umap_tsne_harmony1, fig.width = 8, fig.height = 4, class.source = "fold-hide"}
tsne = Seurat::DimPlot(sobj, group.by = "project_name",
                       reduction = paste0("harmony_", ndims, "_tsne")) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name) +
  Seurat::NoAxes() + ggplot2::ggtitle("PCA - harmony - tSNE") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

umap = Seurat::DimPlot(sobj, group.by = "project_name",
                       reduction = paste0("harmony_", ndims, "_umap")) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name) +
  Seurat::NoAxes() + ggplot2::ggtitle("PCA - harmony - UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

tsne | umap
```

We will keep the UMAP from harmony :

```{r set_name2D}
reduction = "harmony"
name2D = paste0("harmony_", ndims, "_umap")
```

## Clustering

We generate a clustering :

```{r clustering, fig.width = 6, fig.height = 4}
sobj = Seurat::FindNeighbors(sobj, reduction = reduction, dims = 1:ndims)
sobj = Seurat::FindClusters(sobj, resolution = 1)

Seurat::DimPlot(sobj, reduction = name2D, label = TRUE) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
```

# Visualization

We can represent the 4 quality metrics :

```{r qc_plot, fig.width = 12, fig.height = 3}
plot_list = Seurat::FeaturePlot(sobj, reduction = name2D,
                                combine = FALSE, pt.size = 1,
                                features = c("percent.mt", "percent.rb", "percent.st", "nFeature_RNA"))
plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot +
    Seurat::NoAxes() +
    ggplot2::scale_color_gradientn(colors = aquarius:::color_gene) +
    ggplot2::theme(aspect.ratio = 1)
})

patchwork::wrap_plots(plot_list, nrow = 1)
```

We make a representation split by origin to show cell types :

```{r cell_type_split, fig.width = 12, fig.height = 12}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "sample_identifier",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name))

patchwork::wrap_plots(plot_list, ncol = 4) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::guides(color = guide_legend(override.aes = list(size = 4)))
```
## Clusters

We can represent clusters, split by sample of origin :

```{r plot_split_dimred, fig.width = 12, fig.height = 12}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "sample_identifier",
                                        group_by = "seurat_clusters",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = aquarius::gg_color_hue(length(levels(sobj$seurat_clusters))))

patchwork::wrap_plots(plot_list, ncol = 4) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right") &
  guides(color = guide_legend(override.aes = list(size = 3), title = "Cluster"))
```
## Cell type

We annotate cells for cell type using `Seurat::AddModuleScore` function.

```{r cell_annot_custom_short}
sobj = aquarius::cell_annot_custom(sobj,
                                   newname = "cell_type",
                                   markers = cell_markers,
                                   use_negative = TRUE,
                                   add_score = TRUE,
                                   verbose = TRUE)

colnames(sobj@meta.data) = stringr::str_replace_all(string = colnames(sobj@meta.data),
                                                    pattern = " ",
                                                    replacement = "_")

sobj$cell_type = factor(sobj$cell_type, levels = names(cell_markers))

table(sobj$cell_type) 
```

## Cluster smoothing

```{r clustering_smoothing}
cluster_type = table(sobj$cell_type, sobj$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type = setNames(nm = names(cluster_type),
                        levels(sobj$cell_type)[cluster_type])

sobj$cluster_type = cluster_type[sobj$seurat_clusters]
```


We make a table to show the number of cells in each dataset :

```{r table_cell_type_test, fig.width = 11, fig.height = 5, class.source = "fold-hide"}
df_quantities = table(sobj$cell_type,
                      sobj$project_name) %>%
  as.data.frame.array(stringAsFactors = FALSE) %>%
  dplyr::mutate(cell_type = rownames(.)) %>%
  dplyr::relocate(cell_type)
df_quantities$nb_cells = rowSums(table(sobj$cell_type, sobj$project_name))
df_quantities = rbind(df_quantities,
                      c("Number of cells", as.numeric(table(sobj$project_name)), ncol(sobj)))

columns_color = unlist(color_markers[levels(sobj$cell_type)])
columns_color = c(columns_color, NA, rep(NA, times = nrow(df_quantities)*(ncol(df_quantities) - 1)))
font_color = ifelse(is.na(columns_color), yes = "black", no = "white")

# Cell type
custom_theme1 = gridExtra::ttheme_minimal(base_colour = font_color,
                                          core = list(bg_params = list(fill = columns_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb1 = gridExtra::tableGrob(df_quantities,
                           rows = NULL,
                           cols = c("Cell type", levels(sobj$project_name), "Nb cells"),
                           theme = custom_theme1)

# Patchwork
title_gg = ggplot2::ggplot() +
  ggplot2::theme_minimal()

patchwork::wrap_plots(list(title_gg, gb1)) +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15))
```

We make the same with proportions :

```{r prop_cell_type, fig.width = 11, fig.height = 5, class.source = "fold-hide"}
df_quantities = table(sobj$cluster_type,
                      sobj$project_name) %>%
  prop.table(., margin = 2) %>%
  "*"(100) %>% round(., 2) %>%
  as.data.frame.array(stringAsFactors = FALSE) %>%
  dplyr::mutate(cluster_type = rownames(.)) %>%
  dplyr::relocate(cluster_type)
df_quantities$nb_cells = rowSums(table(sobj$cluster_type, sobj$project_name))
df_quantities = rbind(df_quantities,
                      c("Number of cells", as.numeric(table(sobj$project_name)), ncol(sobj)))

columns_color = unlist(color_markers[levels(sobj$cluster_type)])
columns_color = c(columns_color, NA, rep(NA, times = nrow(df_quantities)*(ncol(df_quantities) - 1)))
font_color = ifelse(is.na(columns_color), yes = "black", no = "white")

# Cell type
custom_theme1 = gridExtra::ttheme_minimal(base_colour = font_color,
                                          core = list(bg_params = list(fill = columns_color,
                                                                       col = "black")),
                                          colhead = list(fg_params = list(col = "black")))
gb1 = gridExtra::tableGrob(df_quantities,
                           rows = NULL,
                           cols = c("Cell type", levels(sobj$project_name), "Nb cells"),
                           theme = custom_theme1)

# Patchwork
title_gg = ggplot2::ggplot() +
  ggplot2::theme_minimal()

patchwork::wrap_plots(list(title_gg, gb1)) +
  patchwork::plot_layout(design = "A\nB", heights = c(0.1,5)) &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 15))
```

```{r barplot_stack, fig.width = 10, fig.height = 5}
aquarius::plot_barplot(df = table(sobj$project_name,
                                  sobj$cluster_type) %>%
                         as.data.frame.table() %>%
                         `colnames<-`(c("Sample", "Cell Type", "Number")),
                       x = "Sample", y = "Number", fill = "Cell Type",
                       position = position_fill()) +
  ggplot2::scale_fill_manual(values = unlist(color_markers),
                             breaks = names(color_markers),
                             name = "Cell Type")
```

```{r scores_of_interest}
Seurat::Idents(sobj) = sobj$cluster_type

all_genes = sobj@assays[["RNA"]]@counts %>% rownames()
Col_genes = grep(all_genes, pattern = "^Col", value = TRUE)

# !! Some genes have to be manually removed from the lists first !!

sobj$Col_score = Seurat::AddModuleScore(sobj, features = list(Col_genes))$Cluster1

```


# Save

We save the Seurat object :

```{r save_sobj}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_sobj.rds"))
saveRDS(sample_info, file = paste0(out_dir, "/", save_name, "_sample_info.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

