---
title: "Analysis - immune cells - microenvironment datasets"
subtitle: "Immune cell annotation"
author: "Pernelle"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

```{r clean, echo = FALSE}
rm(list = ls())
```

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <= Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(org.Mm.eg.db)
library(ComplexHeatmap)
```


# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "IC"
out_dir = "."
```


We load the dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/../5_immune_extraction/","immune_sobj.rds"))
sobj
```

and sample information :

```{r load_sample_info}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "env_metadata_sample_info.rds"))
sample_info
```

This is the projection of interest :

```{r name2D}
name2D = "harmony_47_umap"
```

We visualize clusters :


```{r clustering, fig.width = 6, fig.height = 4, eval = FALSE}
Seurat::DimPlot(sobj, reduction = name2D, label = TRUE, pt.size=0.1) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
```

We change the resolution of the dataset :

```{r resolution}
sobj = Seurat::FindClusters(sobj, resolution = 0.5)

Seurat::DimPlot(sobj, reduction = name2D, group.by ="seurat_clusters", label = TRUE, pt.size = 1, label.size = 15) +
  Seurat::NoAxes() +
  Seurat::NoLegend() +
  ggplot2::theme(aspect.ratio = 0.8)
```

```{r cell_annot_IC}
IC_markers = list("T cells" = c("Cd3g", "Cd3d", "Cd3e", "Cd8a", "Cd4", "Foxp3"),
                  "Macrophages" = c("Mrc1", "Pf4", "Apoe","Gas6","Adgre1"),
                  "Dendritic cells" = c("Plbd1","Cd209a", "Mgl2","Cd74"),
                  "Mast cells" = c( "Kit"),
                  "Langerhans cells" = c("Mfge8","Cd207","H2-M2","Epcam","Pxdc1","Cd74"),
                  "NK cells" = c("Nkg7", "Klrk1", "Klri2", "Klrc2", "Klrb1c", "Ncr1", "Klrb1f"))

# Not found in slot data:
# "Neutrophils" = c("S100a8", "S100a9","Retnlg"))

length(IC_markers)

color_markers_IC = aquarius::c30_palette[c(1:length(IC_markers))]
names(color_markers_IC) = names(IC_markers)

data.frame(IC_type = names(color_markers_IC),
           color = unlist(color_markers_IC)) %>%
  ggplot2::ggplot(., aes(x = IC_type, y = 0, fill = IC_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers_IC), breaks = names(color_markers_IC)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank(),
                 axis.text.x = element_text(angle = 10))

sobj = aquarius::cell_annot_custom(sobj,
                                   newname = "IC_type",
                                   markers = IC_markers,
                                   use_negative = TRUE,
                                   add_score = TRUE,
                                   verbose = TRUE)

colnames(sobj@meta.data) = stringr::str_replace_all(string = colnames(sobj@meta.data),
                                                    pattern = " ",
                                                    replacement = "_")

sobj$IC_type = factor(sobj$IC_type, levels = names(IC_markers))

Seurat::DimPlot(sobj, group.by = "IC_type", 
                reduction = name2D,
                cols = color_markers_IC, na.value = "lightgray") +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))
```

## Cluster smoothing

```{r clustering_smoothing}
IC_cluster_type = table(sobj$IC_type, sobj$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type = setNames(nm = names(IC_cluster_type),
                        levels(sobj$IC_type)[IC_cluster_type])

sobj$IC_cluster_type = cluster_type[sobj$seurat_clusters]
```

## New clusters

We visualize the clustering :

```{r see_new.clusters, fig.width = 12, fig.height = 6, class.source = "fold-hide"}
umap = Seurat::DimPlot(sobj, group.by = "IC_cluster_type", label = FALSE,
                       reduction = name2D, cols = color_markers_IC) +
  Seurat::NoAxes() + ggplot2::ggtitle("UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

umap
```

and split by origin :

```{r plot_split_dimred, fig.width = 12, fig.height = 10}
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365"))

plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "sample_type",
                                        group_by = "IC_cluster_type",
                                        main_pt_size = 0.1,
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = color_markers_IC)

patchwork::wrap_plots(plot_list, ncol = 3) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right") &
  guides(color = guide_legend(override.aes = list(size = 3), title = "Cluster"))
```



```{r barplot_count_beautiful, fig.width = 11.77, fig.height = 10}
quantif = table(sobj$sample_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Sample", "nb_cells"))

plot_list = table(sobj$sample_type, sobj$IC_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c( "Sample", "Cell Type", "Number")) %>%
  aquarius::plot_barplot(.,
                         x = "Sample", y = "Number", fill = "Cell Type",
                         position = position_fill()) +
  ggplot2::scale_fill_manual(values = unlist(color_markers_IC),
                             breaks = names(color_markers_IC),
                             name = "Cell Type") +
  ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.5),
                              labels = paste0(seq(0, 100, by = 50)," %"),
                              expand = ggplot2::expansion(add = c(0, 0.05))) +
  ggplot2::geom_label(data = quantif, inherit.aes = FALSE,
                      aes(x = .data$Sample, y = 1.05, label = .data$nb_cells),
                      label.size = 0, size = 8) +
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.line.x = element_line(colour = "lightgray"),
                 text = element_text(size = 15),
                 legend.position = "none")
plot_list
```

```{r barplot_count_beautiful_stack, fig.width = 10, fig.height = 10}
quantif = table(sobj$sample_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Sample", "nb_cells"))

plot_list = table(sobj$sample_type, sobj$IC_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c( "Sample", "Cell Type", "Number")) %>%
  aquarius::plot_barplot(.,
                         x = "Sample", y = "Number", fill = "Cell Type",
                         position = position_stack()) +
  ggplot2::scale_fill_manual(values = unlist(color_markers_IC),
                             breaks = names(color_markers_IC),
                             name = "Cell Type") +
  ggplot2::scale_y_continuous(breaks = seq(0, 1500, by = 200),
                              labels = paste0(seq(0, 1500, by = 200),""),
                              expand = ggplot2::expansion(add = c(0, 30))) +
  ggplot2::geom_label(data = quantif, inherit.aes = FALSE,
                      aes(x = .data$Sample, y = .data$nb_cells+20, label = .data$nb_cells),
                      label.size = 0, size = 8) +
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.line.x = element_line(colour = "lightgray"),
                 text = element_text(size = 15),
                 legend.position = "none")
plot_list
```

```{r check_markers, fig.width = 22, fig.height = 7}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        color_by = "Tgfb1",
                                        split_by = "sample_type",
                                        bg_pt_size = 0.5,
                                        main_pt_size = 0.5)

plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot = one_plot +
    ggplot2::scale_color_gradientn(colors = aquarius::color_gene,
                                 name = "Expression level") +
  # ggplot2::theme(legend.key.height = unit(5,"mm"),
  #                legend.key.width = unit(5,"mm"),
  #                legend.title = element_text(size = 20),
  #                legend.text = element_text(size = 20)) +
  ggplot2::theme(aspect.ratio = 0.8)
  return(one_plot)
})

names(plot_list) = levels(sobj$sample_type)
plot_list = plot_list[c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365")]
patchwork::wrap_plots(plot_list, ncol = 3)
```


```{r dotplot_other_nmSCs_cluster, fig.width = 25, fig.height = 7.5}
Seurat::Idents(sobj) = sobj$IC_cluster_type
sobj$IC_cluster_type = factor(sobj$IC_cluster_type, levels = c("T cells","NK cells","Mast cells","Dendritic cells","Macrophages","Langerhans cells"))

Seurat::DotPlot(sobj,
                features = c("Cd3e","Klrk1","Nkg7","Ncr1","Kit","Cd209a","Lyz2","Adgre1","Epcam","Cd207"),
                group.by = "IC_cluster_type", dot.scale = 15) +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  ggplot2::theme(axis.title = element_blank(),
                 legend.key.height = unit(10,"mm"),
                 legend.key.width = unit(10,"mm"),
                 legend.title = element_text(size = 25),
                 legend.text = element_text(size = 25),
                 axis.text = element_text(size = 30),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggplot2::theme(aspect.ratio = 0.3)
```
# Save

We save the Seurat object :

```{r save_sobj}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_sobj.rds"))
saveRDS(sample_info, file = paste0(out_dir, "/", save_name, "_sample_info.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

