---
title: "Dissociation methods 3h vs 18h"
subtitle: "Make metadata - Skin"
author: "Pernelle"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

```{r clean, echo = FALSE}
rm(list = ls())
```

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


This file is used to generate :

* **sample_info** : dataframe containing sample names, sample path, custom colors...
* **color_markers** (short version) : named list to give custom colors to cell types
* **cell_markers** (short version) : named list containing markers for each cell type


```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
```


# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "dissociation_methods_metadata"
out_dir = "/home/aurelien/Bureau/scRNAseq_methods/1_metadata/"
```

# Sample info

We initialize a dataframe containing everything except colors :

```{r init_sample_info}
sample_info = data.frame(
  project_name = c("2023_10","2023_09","2023_12","2023_11"),
  mouse_id = c(rep("28712",4)),
  age = c(rep("P365",4)),
  dissociation = c(rep("3h", 2),
                   rep("18h",2)),
  genotype = c(rep("flfl", 4)),
  condition = c(rep(c("HLS","cNFs"),2)),
  color = NA,
  stringsAsFactors = FALSE) %>%
  dplyr::mutate(sample_type = paste0(genotype, "_", condition, "_", age)) %>%
  dplyr::mutate(unique_id = data.table::rowid(sample_type)) %>%
  dplyr::mutate(sample_identifier = paste0(mouse_id, "_", age, "_", condition, "_",dissociation)) %>%
  dplyr::select(project_name, mouse_id, genotype, age, condition, dissociation, sample_identifier, sample_type, color)

head(sample_info)
```

We add custom colors for each sample :

```{r custom_palette}
list_palette = list(flfl_HLS_P90 = colorRampPalette(c("cadetblue2", "cadetblue4")),
                    flfl_cNFs_P90 = colorRampPalette(c("orange2", "orange4")),
                    ctrl_HS_P90 = colorRampPalette(c("orchid2", "orchid4")),
                    flfl_HLS_P365 = colorRampPalette(c("chartreuse2", "chartreuse4")),
                    flfl_cNFs_P365 = colorRampPalette(c("firebrick2", "firebrick4")),
                    ctrl_HS_P365 = colorRampPalette(c("purple2", "purple4")))

for (sample_type in unique(sample_info$sample_type)) {
  matching_dataset = (sample_info$sample_type == sample_type)
  sample_info[matching_dataset, "color"] = list_palette[[sample_type]](sum(matching_dataset))
}
```

We visualize the colors :

```{r custom_palette_see, fig.width = 10, fig.height = 1}
sample_info = sample_info %>%
  dplyr::arrange(as.character(sample_identifier)) %>%
  dplyr::mutate(project_name = factor(project_name, levels = project_name)) %>%
  dplyr::mutate(sample_identifier = factor(sample_identifier, levels = sample_identifier))

ggplot2::ggplot(sample_info) +
  ggplot2::geom_label(aes(x = project_name, y = 0, label = project_name, fill = project_name)) +
  ggplot2::geom_text(aes(x = project_name, y = 1, label = sample_identifier), angle = 10) +
  ggplot2::lims(y = c(-1, 2)) +
  ggplot2::scale_fill_manual(values = sample_info$color, breaks = sample_info$project_name) +
  ggplot2::theme_void() +
  ggplot2::theme(legend.position = "none")
```

It is possible to make confusion between colors ?

```{r custom_palette_pie, fig.width = 4, fig.height = 4}
graphics::pie(rep(1, nrow(sample_info)),
              col = sample_info$color,
              labels = sample_info$project_name)
```

We save it :

```{r save_sample_info}
saveRDS(sample_info, file = paste0(out_dir, "/", save_name, "_sample_info.rds"))
```


# Cell types

We define markers for each population of interest :

```{r cell_markers, class.source = "fold-hide"}
cell_markers = list(
  "Myelinating schwann cells" = c("Sox10", "Plp1", "S100b", "Lgi4", "Mpz","Pmp22","Mal","Egr2"),
  "Non myelinating schwann cells" = c("Sox10", "Plp1", "S100b", "Lgi4","Gap43","Ngfr","Scn7a"),
  "Immune cells" = c("Ptprc","Xcr1","Cpa3","Tpsab1","Hdc","Tpsb2"),
  "Fibroblasts" = c("Abi3bp","Ramp1","Mylk","Acta2", "Tagln","Grem2","Pdgfra","Dcn" ,"Sparc","Col3a1","Lum","Cyp2f2","Ccl19" , "Gpx3","Gsn","Crabp1","Pappa2","Corin", "Nrg2","Cntn1","Plac9","Antxr1","Clmp"),
  "Mural cells" = c("Acta2","Rgs5", "Notch3", "Gucy1a1", "Mylk", "Gm13889", "Tagln", "Myh11","Des"),
  "Endothelial cells" = c("Fabp4", "Pecam1", "Cldn5", "Ptprb", "Egfl7"),
  "Melanocytes" = c("Mlana", "Dct"),
  "Keratinocytes" = c("Krt27","Krt28","Krt10","Krt15","Krt14","Krt1","Epcam","Krt5","Cdh1","Gm15987","Abca12"),
  "Muscle cells" = c("Pax7","Myod1","Acta2","Des"))

lengths(cell_markers)
```

Here are custom colors for each cell type :

```{r color_markers, fig.width = 8, fig.height = 0.8, class.source = "fold-hide"}
color_markers = aquarius::c30_palette[c(1:length(cell_markers))]
names(color_markers) = names(cell_markers)

data.frame(cell_type = names(color_markers),
           color = unlist(color_markers)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank(),
                 axis.text.x = element_text(angle = 10))
```


We save them :

```{r save_short_annotation}
saveRDS(color_markers, file = paste0(out_dir, "/", save_name, "_color_markers.rds"))
saveRDS(cell_markers, file = paste0(out_dir, "/", save_name, "_cell_markers.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```
