---
title: "Analysis - vascular cells - microenvironment datasets"
subtitle: "Annotation of vascular cells"
author: "Pernelle"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

```{r clean, echo = FALSE}
rm(list = ls())
```

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <= Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(org.Mm.eg.db)
library(ComplexHeatmap)
```


# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "VC"
out_dir = "."
```


We load the dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/../5_vascular_extraction/","VC_sobj.rds"))
sobj
```

and sample information :

```{r load_sample_info}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "env_metadata_sample_info.rds"))
sample_info
```

This is the projection of interest :

```{r name2D}
name2D = "harmony_43_umap"
```

We visualize clusters :


```{r clustering, fig.width = 6, fig.height = 4}
Seurat::DimPlot(sobj, reduction = name2D, label = TRUE, pt.size=0.1) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
```

```{r cell_annot_VC}
VC_markers = list("Pericytes" = c("Rgs5","Kcnj8","Pdgfrb"),
                  "Vascular endothelial cells" = c("Pecam1","Egfl7","Fabp4","Aqp1","Cdh5"),
                  "Lymphatic endothelial cells" = c("Lyve1","Prox1","Mmrn1","Fgl2","Fxyd6"),
                  "Others" = c("Mfge8","Cd207","H2-M2","Epcam","Pxdc1","Cd74"))

# Not found in slot data:
# "Neutrophils" = c("S100a8", "S100a9","Retnlg"))

length(VC_markers)
color_markers_VC = c("blue4","firebrick4","seagreen4","orchid3")


names(color_markers_VC) = names(VC_markers)

data.frame(VC_type = names(color_markers_VC),
           color = unlist(color_markers_VC)) %>%
  ggplot2::ggplot(., aes(x = VC_type, y = 0, fill = VC_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers_VC), breaks = names(color_markers_VC)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank(),
                 axis.text.x = element_text(angle = 10))

sobj = aquarius::cell_annot_custom(sobj,
                                   newname = "VC_type",
                                   markers = VC_markers,
                                   use_negative = TRUE,
                                   add_score = TRUE,
                                   verbose = TRUE)

colnames(sobj@meta.data) = stringr::str_replace_all(string = colnames(sobj@meta.data),
                                                    pattern = " ",
                                                    replacement = "_")

sobj$VC_type = factor(sobj$VC_type, levels = names(VC_markers))

Seurat::DimPlot(sobj, group.by = "VC_type", 
                reduction = name2D,
                cols = color_markers_VC,
                na.value = "lightgray") +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))
```

## Cluster smoothing

```{r clustering_smoothing}
VC_cluster_type = table(sobj$VC_type, sobj$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type = setNames(nm = names(VC_cluster_type),
                        levels(sobj$VC_type)[VC_cluster_type])

sobj$VC_cluster_type = cluster_type[sobj$seurat_clusters]

sobj$VC_cluster_type = factor(sobj$VC_cluster_type, levels = names(VC_markers))
```

## New clusters

We visualize the clustering :

```{r see_new.clusters, fig.width = 12, fig.height = 6, class.source = "fold-hide"}
umap = Seurat::DimPlot(sobj, group.by = "VC_cluster_type", label = FALSE,
                       reduction = name2D, cols = color_markers_VC) +
  Seurat::NoAxes() + ggplot2::ggtitle("UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))
umap
```

We visualize the clustering :

```{r remove_other_cells, fig.width = 12, fig.height = 6, class.source = "fold-hide"}

Seurat::Idents(sobj) = sobj$VC_cluster_type
sobj = subset(sobj, invert = TRUE, idents = "Others")


umap = Seurat::DimPlot(sobj, group.by = "VC_cluster_type", label = FALSE,
                       reduction = name2D, cols = color_markers_VC) +
  Seurat::NoAxes() + ggplot2::ggtitle("UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))
umap
```

and split by origin :

```{r plot_split_dimred, fig.width = 12, fig.height = 10}
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365"))

plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "sample_type",
                                        group_by = "VC_cluster_type",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = color_markers_VC)

patchwork::wrap_plots(plot_list, ncol = 3) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right") &
  guides(color = guide_legend(override.aes = list(size = 3), title = "Cluster"))
```



```{r barplot_count_beautiful_prep}
quantif = table(sobj$sample_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Sample", "nb_cells"))

quantif2 = table(sobj$VC_cluster_type, sobj$sample_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Cell_type", "Sample_type","nb_cell")) %>%
  dplyr::filter(Cell_type != "Others") %>%
  dplyr::mutate(Cell_type = as.character(Cell_type)) %>%
  dplyr::group_by(Sample_type) %>%
  dplyr::mutate(prop_cell = nb_cell/sum(nb_cell)) %>%
  dplyr::mutate(prop_cum =  cumsum(nb_cell/sum(nb_cell))) %>%
  dplyr::mutate(prop_cum_mid = prop_cum) %>%
  as.data.frame()

quantif3 = lapply(unique(quantif2$Sample_type), FUN = function(Sample_type) {
  subquantif2 = quantif2[quantif2$Sample_type == Sample_type, ]
  
  for (i in c(1:nrow(subquantif2))) {
    if (i == 1) {
      subquantif2[i, "prop_cum_mid"] = subquantif2[i, "prop_cum"]/2
    } else {
      subquantif2[i, "prop_cum_mid"] = (subquantif2[i - 1, "prop_cum"] + subquantif2[i, "prop_cum"])/2
    }
  }
  
  return(subquantif2)
}) %>% do.call(rbind.data.frame, .) %>%
  dplyr::mutate(prop_cell = 100*round(prop_cell, 4))
```





```{r barplot_count_beautiful, fig.width = 11.77, fig.height = 10}
table(sobj$sample_type, sobj$VC_cluster_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c( "Sample", "Cell Type", "Number")) %>%
  aquarius::plot_barplot(.,
                         x = "Sample", y = "Number", fill = "Cell Type",
                         position = position_fill()) +
  # Fill color
  ggplot2::scale_fill_manual(values = unlist(color_markers_VC),
                             breaks = names(color_markers_VC),
                             name = "Cell Type") +
  # Y axis style
  ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.5),
                              labels = paste0(seq(0, 100, by = 50)," %"),
                              expand = ggplot2::expansion(add = c(0, 0.05))) +
  # Number of cells by sample
  ggplot2::geom_label(data = quantif, inherit.aes = FALSE,
                      aes(x = Sample, y = 1.03, label = nb_cells),
                      label.size = 0, size = 5) +
  # Proportion of cell types by sample
  ggplot2::geom_label(data = quantif3,
                      inherit.aes = FALSE, fill = "white",
                      label.size = 0, size = 5,
                      aes(x = Sample_type,
                          y = 1 - prop_cum_mid,
                          label = paste0(prop_cell, "%"))) +
  # Global style
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.line.x = element_line(colour = "lightgray"),
                 text = element_text(size = 15),
                 legend.position = "none")
  
```



# Save

We save the Seurat object :

```{r save_sobj}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_sobj.rds"))
saveRDS(sample_info, file = paste0(out_dir, "/", save_name, "_sample_info.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

