---
params:
  sample_name: " "
title: "Microenvironment cells"
subtitle: "Sample `r params$sample_name`"
author: "Pernelle"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # useful for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 10,
                      fig.height = 10,
                      
                      # add runtime after chunk
                      time_it = FALSE)
```

```{r sample_name}
#sample_name = "2023_20"
sample_name = params$sample_name
```


The goal of this script is to generate a Seurat object for sample `r sample_name`.

* removal of cells with less than 600 genes or less than 6 log UMI
* normalization with `LogNormalize`, then doublets detection using `scran hybrid` and `scDblFinder` method, and doublet cells removal
* removal of cells expressing more than 20 % of mitochondrial genes, or more than 30 % of ribosomal genes
* normalization with `LogNormalize`, for only the remaining cells
* cell cycle and cell type annotation
* dimensionality reduction using `PCA`
* projection using `tSNE` and `UMAP`


```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(BiocParallel)
library(Seurat)
.libPaths()
```


# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
out_dir = "."
```

Input count matrix is there :

```{r count_matrix_dir}
count_matrix_dir = paste0(out_dir, "/../0_input/", sample_name)
```

These is a parameter for different functions :

```{r global_settings}
cl = BiocParallel::SerialParam()
#aquarius::create_parallel_instance(nthreads = 1L)
cut_log_nCount_RNA = 6
cut_nFeature_RNA = 600
cut_percent.mt = 0.2
cut_percent.rb = 0.4
```

We load metadata :

```{r load_metadata}
color_markers = readRDS(paste0(out_dir, "/../1_metadata/", "env_metadata_color_markers.rds"))
cell_markers = readRDS(paste0(out_dir, "/../1_metadata/", "env_metadata_cell_markers.rds"))
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "env_metadata_sample_info.rds"))
```

# Load count matrix

In this section, we load the raw count matrix. Then, we applied an empty droplets filtering.

```{r combine_datasets}
sobj = aquarius::load_sc_data(data_path = count_matrix_dir,
                              sample_name = sample_name,
                              my_seed = 1337L)
sobj
```

We add the same columns as in metadata :

```{r add_metadata}
row_oi = (sample_info$project_name == sample_name)

sobj$project_name = sample_name
sobj$exp_date = sample_info[row_oi, "exp_date"]
sobj$age = sample_info[row_oi, "age"]
sobj$genotype = sample_info[row_oi, "genotype"]
sobj$condition = sample_info[row_oi, "condition"]
sobj$sample_identifier = sample_info[row_oi, "sample_identifier"]
sobj$sample_type = sample_info[row_oi, "sample_type"]

colnames(sobj@meta.data)
```

# Before filtering

## Gene expression normalization

```{r normalization}
sobj = Seurat::NormalizeData(sobj,
                             normalization.method = "LogNormalize",
                             assay = "RNA")

sobj = Seurat::FindVariableFeatures(sobj,
                                    assay = "RNA",
                                    nfeatures = 3000)
sobj
```

## Projection

We generate a tSNE to visualize cells before filtering.

```{r pca_before, fig.width = 12, fig.height = 4}
sobj = aquarius::dimensions_reduction(sobj = sobj,
                                      assay = "RNA",
                                      reduction = "pca",
                                      max_dims = 100,
                                      verbose = FALSE)
Seurat::ElbowPlot(sobj, ndims = 100, reduction = "RNA_pca")
```
We choose the number of dimensions such that they summarize 60 % of the variability :

```{r ndims}
ndims = 20
```

We generate a tSNE with ndims principal components :

```{r tsne_before}
sobj = Seurat::RunTSNE(sobj,
                       reduction = "RNA_pca",
                       dims = 1:ndims,
                       seed.use = 1337L,
                       reduction.name = paste0("RNA_pca_", ndims, "_tsne"))

sobj
```


## Cell type

We annotate cells for cell type using `Seurat::AddModuleScore` function.

```{r cell_annot_custom_short}
sobj = aquarius::cell_annot_custom(sobj,
                                   newname = "cell_type",
                                   markers = cell_markers,
                                   use_negative = TRUE,
                                   add_score = TRUE,
                                   verbose = TRUE)

colnames(sobj@meta.data) = stringr::str_replace_all(string = colnames(sobj@meta.data),
                                                    pattern = " ",
                                                    replacement = "_")

sobj$cell_type = factor(sobj$cell_type, levels = names(cell_markers))

table(sobj$cell_type) 
```


To justify cell type annotation, we can make a dotplot :

```{r dotplot_cell_type_short, fig.width = 12, fig.height = 9}
markers = unique(unlist(cell_markers[levels(sobj$cell_type)]))
markers = markers[markers %in% rownames(sobj)]
markers = unique(markers)
aquarius::plot_dotplot(sobj, assay = "RNA",
                       column_name = "cell_type",
                       markers = c(markers),
                       nb_hline = 0) +
  ggplot2::scale_color_gradientn(colors = aquarius:::color_gene) +
  ggplot2::labs(x = "Cell type", y = "") +
  ggplot2::theme(legend.position = "right",
                 legend.box = "vertical",
                 legend.direction = "vertical")
```

We can make a barplot to see the composition of each dataset, and visualize cell types on the projection.

```{r barplot_celltype, fig.height = 6, fig.width = 8, class.source = "fold-hide"}
df_proportion = as.data.frame(prop.table(table(sobj$orig.ident,
                                               sobj$cell_type)))
colnames(df_proportion) = c("orig.ident", "cell_type", "freq")

quantif = table(sobj$orig.ident) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("orig.ident", "nb_cells"))

# Plot
plot_list = list()

plot_list[[2]] = aquarius::plot_barplot(df = df_proportion,
                                        x = "orig.ident",
                                        y = "freq",
                                        fill = "cell_type",
                                        position = ggplot2::position_fill()) +
  ggplot2::scale_fill_manual(name = "Cell type",
                             values = color_markers[levels(df_proportion$cell_type)],
                             breaks = levels(df_proportion$cell_type)) +
  ggplot2::geom_label(data = quantif, inherit.aes = FALSE,
                      aes(x = orig.ident, y = 1.05, label = nb_cells),
                      label.size = 0)

plot_list[[1]] = Seurat::DimPlot(sobj, group.by = "cell_type") +
  ggplot2::scale_color_manual(values = unlist(color_markers),
                              breaks = names(color_markers)) +
  ggplot2::labs(title = sample_name,
                subtitle = paste0(ncol(sobj), " cells")) +
  Seurat::NoLegend() + Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 plot.subtitle = element_text(hjust = 0.5))

patchwork::wrap_plots(plot_list, nrow = 1, widths = c(6, 1))
```

## Cell cycle phase

We annotate cells for cell cycle phase using `Seurat` and `cyclone`.

```{r cell_cycle}
cc_columns = aquarius::add_cell_cycle(sobj = sobj,
                                      assay = "RNA",
                                      species_rdx = "mm",
                                      BPPARAM = cl)@meta.data[, c("Seurat.Phase", "Phase")]

sobj$Seurat.Phase = cc_columns$Seurat.Phase
sobj$cyclone.Phase = cc_columns$Phase

table(sobj$Seurat.Phase, sobj$cyclone.Phase)
```

We visualize cell cycle on the projection :

```{r see_cell_cycle1, fig.width = 12, fig.height = 7, class.source = "fold-hide"}
plot_list = list()

plot_list[[2]] = Seurat::DimPlot(sobj, group.by = "Seurat.Phase") +
  ggplot2::labs(title = "Cell Cycle Phase",
                subtitle = "Seurat.Phase") +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 plot.subtitle = element_text(hjust = 0.5))

plot_list[[1]] = Seurat::DimPlot(sobj, group.by = "cyclone.Phase") +
  ggplot2::labs(title = "Cell Cycle Phase",
                subtitle = "cyclone.Phase") +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 plot.subtitle = element_text(hjust = 0.5))

patchwork::wrap_plots(plot_list, nrow = 1)
```

## Quality control

In this section, we look at the number of genes expressed by each cell, the number of UMI, the percentage of mitochondrial genes expressed, and the percentage of ribosomal genes expressed. Then, without taking into account the cells expressing low number of genes or have low number of UMI, we identify doublet cells.

We compute four quality metrics :

```{r qc_metrics, results = "hide"}
sobj = aquarius::add_QC_metrics(sobj = sobj,
                                species_rdx = "mm",
                                BPPARAM = cl)
```

We get the cell barcodes for the failing cells :

```{r failed}
fail_percent.mt = sobj@meta.data %>% dplyr::filter(percent.mt > cut_percent.mt) %>% rownames()
fail_percent.rb = sobj@meta.data %>% dplyr::filter(percent.rb > cut_percent.rb) %>% rownames()
fail_log_nCount_RNA = sobj@meta.data %>% dplyr::filter(log_nCount_RNA < cut_log_nCount_RNA) %>% rownames()
fail_nFeature_RNA = sobj@meta.data %>% dplyr::filter(nFeature_RNA < cut_nFeature_RNA) %>% rownames()
```

```{r save_temp_sobj}
saveRDS(sobj, file = paste0(out_dir, "/datasets/", sample_name, "_sobj_temp.rds"))
```

## Doublet cells detection

Without taking into account the low UMI and low number of features cells, we identify doublets.

```{r fsobj}
fsobj = subset(sobj, invert = TRUE,
               cells = unique(c(fail_log_nCount_RNA, fail_nFeature_RNA)))
fsobj
```

On this filtered dataset, we apply doublet cells detection. Just before, we run the normalization, taking into account only the remaining cells.

```{r normalization_doublets}
fsobj = aquarius::sc_normalization(sobj = fsobj,
                                   assay = "RNA",
                                   method = "LogNormalize",
                                   verbose = FALSE)
```

We identify doublet cells :

```{r doublet_detection}
fsobj = aquarius::find_doublets(sobj = fsobj,
                                BPPARAM = cl)

fail_doublets_consensus = Seurat::WhichCells(fsobj, expression = doublets_consensus.class)
fail_doublets_scDblFinder = Seurat::WhichCells(fsobj, expression = scDblFinder.class)
fail_doublets_hybrid = Seurat::WhichCells(fsobj, expression = hybrid_score.class)
```

We add the information in the non filtered Seurat object :

```{r add_doublet_metrics}
sobj$doublets_consensus.class = dplyr::case_when(!(colnames(sobj) %in% colnames(fsobj)) ~ NA,
                                                 colnames(sobj) %in% fail_doublets_consensus ~ TRUE,
                                                 !(colnames(sobj) %in% fail_doublets_consensus) ~ FALSE)

sobj$scDblFinder.class = dplyr::case_when(!(colnames(sobj) %in% colnames(fsobj)) ~ NA,
                                          colnames(sobj) %in% fail_doublets_scDblFinder ~ TRUE,
                                          !(colnames(sobj) %in% fail_doublets_scDblFinder) ~ FALSE)

sobj$hybrid_score.class = dplyr::case_when(!(colnames(sobj) %in% colnames(fsobj)) ~ NA,
                                           colnames(sobj) %in% fail_doublets_hybrid ~ TRUE,
                                           !(colnames(sobj) %in% fail_doublets_hybrid) ~ FALSE)
```

```{r clean_fsobj, echo = FALSE}
rm(fsobj)
```


## Quality control representation

We can visualize the 4 cells quality with a Venn diagram : 

```{r qc_venn, class.source = "fold-hide", fig.width = 8, fig.height = 6}
ggvenn::ggvenn(list(percent.mt = fail_percent.mt,
                    percent.rb = fail_percent.rb,
                    log_nCount_RNA = fail_log_nCount_RNA,
                    nFeature_RNA = fail_nFeature_RNA), 
               fill_color = c("#0073C2FF", "#EFC000FF", "orange", "pink"),
               stroke_size = 0.5, set_name_size = 4) +
  ggplot2::ggtitle(label = "Filtered out cells") +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```


### Number of UMI

To visualize the threshold for number of UMI, we can make a histogram :

```{r qc_umi_hist, class.source = "fold-hide", fig.width = 10, fig.height = 5}
aquarius::plot_qc_density(df = sobj@meta.data,
                          x = "log_nCount_RNA",
                          bins = 200,
                          group_by = "orig.ident",
                          group_color = setNames(sample_info$color,
                                                 nm = sample_info$sample_identifiant),
                          x_thresh = cut_log_nCount_RNA) +
  ggplot2::lims(x = c(0, max(sobj$log_nCount_RNA)))
```

```{r vln_umi_cell_type, fig.width = 12, fig.height = 6}
Seurat::VlnPlot(sobj, features = "log_nCount_RNA", pt.size = 0.001,
                group.by = "cell_type", cols = color_markers) +
  ggplot2::scale_fill_manual(values = color_markers, breaks = names(color_markers)) +
  ggplot2::geom_hline(yintercept = cut_log_nCount_RNA, col = "red") +
  ggplot2::labs(x = "") +
  ggplot2::lims(y = c(0, max(sobj$log_nCount_RNA)))
```


### Number of features

To visualize the threshold for number of features, we can make a histogram :

```{r qc_features_hist, class.source = "fold-hide", fig.width = 10, fig.height = 5}
aquarius::plot_qc_density(df = sobj@meta.data,
                          x = "nFeature_RNA",
                          bins = 200,
                          group_by = "orig.ident",
                          group_color = setNames(sample_info$color,
                                                 nm = sample_info$sample_identifiant),
                          x_thresh = cut_nFeature_RNA)
```


```{r vln_features_cell_type, fig.width = 12, fig.height = 6}
Seurat::VlnPlot(sobj, features = "nFeature_RNA", pt.size = 0.001,
                group.by = "cell_type", cols = color_markers) +
  ggplot2::scale_fill_manual(values = color_markers, breaks = names(color_markers)) +
  ggplot2::geom_hline(yintercept = cut_nFeature_RNA, col = "red") +
  ggplot2::labs(x = "")
```


### Mitochondrial genes expression

To identify a threshold for mitochondrial gene expression, we can make a histogram :

```{r qc_mito_hist, class.source = "fold-hide", fig.width = 10, fig.height = 5}
aquarius::plot_qc_density(df = sobj@meta.data,
                          x = "percent.mt",
                          bins = 200,
                          group_by = "orig.ident",
                          group_color = setNames(sample_info$color,
                                                 nm = sample_info$sample_identifiant),
                          x_thresh = cut_percent.mt)
```

```{r vln_percentmt_cell_type, fig.width = 12, fig.height = 6}
Seurat::VlnPlot(sobj, features = "percent.mt", pt.size = 0.001,
                group.by = "cell_type", cols = color_markers) +
  ggplot2::scale_fill_manual(values = color_markers, breaks = names(color_markers)) +
  ggplot2::geom_hline(yintercept = cut_percent.mt, col = "red") +
  ggplot2::labs(x = "")
```


### Ribosomal genes expression

To identify a threshold for ribosomal gene expression, we can make a histogram :

```{r qc_ribo_hist, class.source = "fold-hide", fig.width = 10, fig.height = 5}
aquarius::plot_qc_density(df = sobj@meta.data,
                          x = "percent.rb",
                          bins = 200,
                          group_by = "orig.ident",
                          group_color = setNames(sample_info$color,
                                                 nm = sample_info$sample_identifiant),
                          x_thresh = cut_percent.rb)
```


```{r vln_percentrb_cell_type, fig.width = 12, fig.height = 6}
Seurat::VlnPlot(sobj, features = "percent.rb", pt.size = 0.001,
                group.by = "cell_type", cols = color_markers) +
  ggplot2::scale_fill_manual(values = color_markers, breaks = names(color_markers)) +
  ggplot2::geom_hline(yintercept = cut_percent.rb, col = "red") +
  ggplot2::labs(x = "")
```

### FACS-like figure

We would like to see if the number of feature expressed by cell, and the number of UMI is correlated with the cell type, the percentage of mitochondrial and ribosomal gene expressed, and the doublet status. We build the `log_nCount_RNA` by `nFeature_RNA` figure, where cells (dots) are colored by these different metrics.

This is the figure, colored by cell type :

```{r qc_patchwork_cell_type, class.source = "fold-hide", fig.width = 10, fig.height = 8}
aquarius::plot_qc_facslike(df = sobj@meta.data,
                           x = "nFeature_RNA",
                           y = "log_nCount_RNA",
                           col_by = "cell_type",
                           col_colors = unname(color_markers),
                           x_thresh = cut_nFeature_RNA,
                           y_thresh = cut_log_nCount_RNA,
                           bins = 200)
```

This is the figure, colored by the percentage of mitochondrial genes expressed in cell :

```{r qc_patchwork_mito, class.source = "fold-hide", fig.width = 10, fig.height = 8}
aquarius::plot_qc_facslike(df = sobj@meta.data,
                           x = "nFeature_RNA",
                           y = "log_nCount_RNA",
                           col_by = "percent.mt",
                           x_thresh = cut_nFeature_RNA,
                           y_thresh = cut_log_nCount_RNA,
                           bins = 200)
```

This is the figure, colored by the percentage of ribosomal genes expressed in cell :

```{r qc_patchwork_ribo, class.source = "fold-hide", fig.width = 10, fig.height = 8}
aquarius::plot_qc_facslike(df = sobj@meta.data,
                           x = "nFeature_RNA",
                           y = "log_nCount_RNA",
                           col_by = "percent.rb",
                           x_thresh = cut_nFeature_RNA,
                           y_thresh = cut_log_nCount_RNA,
                           bins = 200)
```

This is the figure, colored by the doublet cells status (`doublets_consensus.class`) :

```{r qc_patchwork_doublet_consensus, class.source = "fold-hide", fig.width = 10, fig.height = 8}
aquarius::plot_qc_facslike(df = sobj@meta.data,
                           x = "nFeature_RNA",
                           y = "log_nCount_RNA",
                           col_by = "doublets_consensus.class",
                           x_thresh = cut_nFeature_RNA,
                           y_thresh = cut_log_nCount_RNA,
                           bins = 200)
```

This is the figure, colored by the doublet cells status (`scDblFinder.class`) :

```{r qc_patchwork_scdblfinder, class.source = "fold-hide", fig.width = 10, fig.height = 8}
aquarius::plot_qc_facslike(df = sobj@meta.data,
                           x = "nFeature_RNA",
                           y = "log_nCount_RNA",
                           col_by = "scDblFinder.class",
                           x_thresh = cut_nFeature_RNA,
                           y_thresh = cut_log_nCount_RNA,
                           bins = 200)
```

This is the figure, colored by the doublet cells status (`hybrid_score.class`) :

```{r qc_patchwork_hybrid_score, class.source = "fold-hide", fig.width = 10, fig.height = 8}
aquarius::plot_qc_facslike(df = sobj@meta.data,
                           x = "nFeature_RNA",
                           y = "log_nCount_RNA",
                           col_by = "hybrid_score.class",
                           x_thresh = cut_nFeature_RNA,
                           y_thresh = cut_log_nCount_RNA,
                           bins = 200)
```


### Visualization as piechart

Do filtered cells belong to a particular cell type ?

```{r qc_piechart_cell_type, class.source = "fold-hide", fig.width = 12, fig.height = 9}
sobj$all_cells = TRUE

piechart_all_cells = aquarius::plot_piechart(df = sobj@meta.data,
                                             logical_var = "all_cells",
                                             grouping_var = "cell_type",
                                             colors = color_markers,
                                             display_legend = TRUE) +
  ggplot2::labs(title = "All cells",
                subtitle = paste(nrow(sobj@meta.data), "cells")) +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                 plot.subtitle = element_text(hjust = 0.5))

piechart_doublets = aquarius::plot_piechart(df = sobj@meta.data %>%
                                              dplyr::filter(doublets_consensus.class),
                                            logical_var = "all_cells",
                                            grouping_var = "cell_type",
                                            colors = color_markers,
                                            display_legend = TRUE) +
  ggplot2::labs(title = "doublets_consensus.class",
                subtitle = paste(sum(sobj$doublets_consensus.class, na.rm = TRUE), "cells")) +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                 plot.subtitle = element_text(hjust = 0.5))

piechart_percent.mt = aquarius::plot_piechart(df = sobj@meta.data %>%
                                                dplyr::filter(percent.mt > cut_percent.mt),
                                              logical_var = "all_cells",
                                              grouping_var = "cell_type",
                                              colors = color_markers,
                                              display_legend = TRUE) +
  ggplot2::labs(title = paste("percent.mt >", cut_percent.mt),
                subtitle = paste(length(fail_percent.mt), "cells")) +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                 plot.subtitle = element_text(hjust = 0.5))

# piechart_percent.rb = aquarius::plot_piechart(df = sobj@meta.data %>%
#                                                 dplyr::filter(percent.rb > cut_percent.rb),
#                                               logical_var = "all_cells",
#                                               grouping_var = "cell_type",
#                                               colors = color_markers,
#                                               display_legend = TRUE) +
#   ggplot2::labs(title = paste("percent.rb >", cut_percent.rb),
#                 subtitle = paste(length(fail_percent.rb), "cells")) +
#   ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"),
#                  plot.subtitle = element_text(hjust = 0.5))

piechart_log_nCount_RNA = aquarius::plot_piechart(df = sobj@meta.data %>%
                                                    dplyr::filter(log_nCount_RNA < cut_log_nCount_RNA),
                                                  logical_var = "all_cells",
                                                  grouping_var = "cell_type",
                                                  colors = color_markers,
                                                  display_legend = TRUE) +
  ggplot2::labs(title = paste("log_nCount_RNA <", round(cut_log_nCount_RNA, 2)),
                subtitle = paste(length(fail_log_nCount_RNA), "cells")) +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                 plot.subtitle = element_text(hjust = 0.5))

piechart_nFeature_RNA = aquarius::plot_piechart(df = sobj@meta.data %>%
                                                  dplyr::filter(nFeature_RNA < cut_nFeature_RNA),
                                                logical_var = "all_cells",
                                                grouping_var = "cell_type",
                                                colors = color_markers,
                                                display_legend = TRUE) +
  ggplot2::labs(title = paste("nFeature_RNA <", round(cut_nFeature_RNA, 2)),
                subtitle = paste(length(fail_nFeature_RNA), "cells")) +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"),
                 plot.subtitle = element_text(hjust = 0.5))

patchwork::wrap_plots(piechart_all_cells, piechart_percent.mt,
                      piechart_doublets, piechart_nFeature_RNA, piechart_log_nCount_RNA) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right")
```

```{r clean_qc_4, echo = FALSE}
sobj$all_cells = NULL

rm(piechart_all_cells, piechart_log_nCount_RNA,
   piechart_nFeature_RNA, piechart_percent.mt,
   piechart_doublets, piechart_percent.rb)
```


### Doublet cells status

We can compare doublet detection methods with a Venn diagram : 

```{r qc_venn_doublet, class.source = "fold-hide", fig.width = 8, fig.height = 6}
ggvenn::ggvenn(list(hybrid = fail_doublets_hybrid,
                    scDblFinder = fail_doublets_scDblFinder), 
               fill_color = c("#0073C2FF", "#EFC000FF"),
               stroke_size = 0.5, set_name_size = 4) +
  ggplot2::ggtitle(label = "Doublet cells") +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```


What is the composition of doublet cells ? We could use deconvolution. We will just look at score for each cell type.

```{r doublets_composition, class.source = "fold-hide"}
# sobj$orig.ident.doublets = case_when(is.na(sobj$doublets_consensus.class) ~ "bad quality",
#                                      sobj$doublets_consensus.class == TRUE ~ paste0(sobj$orig.ident, " doublets"),
#                                      sobj$doublets_consensus.class == FALSE ~ "not doublet")
# sobj$orig.ident.doublets = factor(sobj$orig.ident.doublets,
#                                   levels = c(paste0(as.character(sample_info$sample_identifiant), " doublets"),
#                                              "bad quality", "not doublet"))
# 
# doublets_compo = function(score1, score2) {
#   type1 = unlist(stringr::str_remove(score1, pattern = "score_"))
#   type2 = unlist(stringr::str_remove(score2, pattern = "score_"))
#   
#   if (type1 == type2) {
#     the_title = "Homotypic doublet"
#     the_subtitle = type1
#     score1 = "log_nCount_RNA"
#   } else {
#     the_title = "Heterotypic doublet"
#     the_subtitle = paste(type1, type2, sep = " + ")
#   }
#   
#   df = sobj@meta.data[, c(score1, score2, "orig.ident.doublets")]  %>%
#     dplyr::arrange(desc(orig.ident.doublets))
#   colnames(df) = c("score1", "score2", "orig.ident.doublets")
#   
#   p = ggplot2::ggplot(df, aes(x = score1, y = score2, col = orig.ident.doublets)) +
#     ggplot2::geom_point(size = 0.25) +
#     ggplot2::scale_color_manual(values = c(sample_info$color, "gray90", "gray60"),
#                                 breaks = c(paste0(as.character(sample_info$sample_identifiant), " doublets"),
#                                            "bad quality", "not doublet")) +
#     ggplot2::labs(x = score1, y = score2,
#                   title = the_title, subtitle = the_subtitle) +
#     ggplot2::theme_classic() +
#     ggplot2::theme(aspect.ratio = 1,
#                    plot.title = element_text(hjust = 0.5),
#                    plot.subtitle = element_text(hjust = 0.5))
#   
#   return(p)
# }
```

```{r doublets_patchwork_prep, class.source = "fold-hide"}
# score_columns = grep(x = colnames(sobj@meta.data),
#                      pattern = "^score",
#                      value = TRUE)
# combinations = expand.grid(score_columns, score_columns) %>%
#   apply(., 1, sort) %>% t() %>%
#   as.data.frame()
# combinations = combinations[!duplicated(combinations), ]
# 
# plot_list = apply(combinations, 1, FUN = function(elem) {
#   doublets_compo(elem[1], elem[2])
# })
# 
# sobj$orig.ident.doublets = NULL
```


```{r doublets_patchwork, fig.width = 12, fig.height = 80, fold_plot = TRUE}
# patchwork::wrap_plots(plot_list, ncol = 4) +
#   patchwork::plot_layout(guides = "collect") &
#   ggplot2::theme(legend.position = "right")
```

## Save

We save this object before filtering :

```{r save_sobj_unfiltered_annotated}
saveRDS(sobj, file = paste0(out_dir, "/datasets/", sample_name, "_sobj_unfiltered.rds"))
```


# Filtering

We remove :

* cells with a number of UMI lower than `r cut_log_nCount_RNA`
* cells expressing a number of genes lower than `r cut_nFeature_RNA`
* cells expressing more than `r 100*cut_percent.mt` \% of mitochondrial genes
* cells expressing more than `r 100*cut_percent.rb` \% of ribosomal protein coding genes
* doublet cells detected with both method (**scDblFinder** and **scds-hybrid**)


```{r filter_cells}
sobj = subset(sobj, invert = TRUE,
              cells = unique(c(fail_log_nCount_RNA, fail_nFeature_RNA,
                               fail_percent.mt, fail_percent.rb,
                               fail_doublets_consensus)))
sobj
```


```{r save_filtering_metrics, eval = FALSE, echo = FALSE}
save(fail_doublets_consensus, fail_doublets_scDblFinder, fail_doublets_hybrid,
     fail_percent.mt, fail_log_nCount_RNA, fail_nFeature_RNA,
     cut_percent.mt, cut_log_nCount_RNA, cut_nFeature_RNA,
     file = paste0(out_dir, "/", sample_name, "_filtered_cells.rda"))
```

```{r clean_filter, echo = FALSE}
rm(fail_doublets_consensus, fail_doublets_scDblFinder, fail_doublets_hybrid,
   fail_percent.mt, fail_log_nCount_RNA, fail_nFeature_RNA,
   cut_percent.mt, cut_log_nCount_RNA, cut_nFeature_RNA)
```


# Post-filtering processing

We normalize the count matrix for remaining cells :

```{r normalization_3}
sobj = aquarius::sc_normalization(sobj = sobj,
                                  assay = "RNA",
                                  method = "LogNormalize",
                                  verbose = FALSE)
```


We perform a PCA :

```{r pca, fig.width = 12, fig.height = 4}
sobj = aquarius::dimensions_reduction(sobj = sobj,
                                      assay = "RNA",
                                      reduction = "pca",
                                      max_dims = 100,
                                      verbose = FALSE)
Seurat::ElbowPlot(sobj, ndims = 100, reduction = "RNA_pca")
```


We generate a tSNE and a UMAP with 20 principal components :

```{r tsne_umap}
ndims = 20
sobj = Seurat::RunTSNE(sobj,
                       reduction = "RNA_pca",
                       dims = 1:ndims,
                       seed.use = 1337L,
                       reduction.name = paste0("RNA_pca_", ndims, "_tsne"))

sobj = Seurat::RunUMAP(sobj,
                       reduction = "RNA_pca",
                       dims = 1:ndims,
                       seed.use = 1337L,
                       reduction.name = paste0("RNA_pca_", ndims, "_umap"))
```


## Annotation

### Cell type simple

We annotate cells for cell type, with the new normalized expression matrix :

```{r cell_type_2, class.source = "fold-hide"}
score_columns = grep(x = colnames(sobj@meta.data), pattern = "^score", value = TRUE)
sobj@meta.data[, score_columns] = NULL
sobj$cell_type = NULL

sobj = aquarius::cell_annot_custom(sobj,
                                   newname = "cell_type",
                                   markers = cell_markers,
                                   use_negative = TRUE,
                                   add_score = TRUE,
                                   verbose = TRUE)

sobj$cell_type = factor(sobj$cell_type, levels = names(cell_markers))

colnames(sobj@meta.data) = stringr::str_replace_all(string = colnames(sobj@meta.data),
                                                    pattern = " ",
                                                    replacement = "_")

table(sobj$cell_type)
```


To justify cell type annotation, we can make a dotplot :

```{r dotplot_cell_type_short2, fig.width = 12, fig.height = 9}
markers = unique(unlist(cell_markers[levels(sobj$cell_type)]))
markers = markers[markers %in% rownames(sobj)]
markers = unique(markers)

aquarius::plot_dotplot(sobj, assay = "RNA",
                       column_name = "cell_type",
                       markers = c(markers),
                       nb_hline = 0) +
  ggplot2::scale_color_gradientn(colors = aquarius:::color_gene) +
  ggplot2::labs(x = "Cell type", y = "") +
  ggplot2::theme(legend.position = "right",
                 legend.box = "vertical",
                 legend.direction = "vertical")
```

We can make a barplot to see the composition of each dataset.

```{r barplot_celltype2, fig.height = 5, fig.width = 5}
df_proportion = as.data.frame(prop.table(table(sobj$orig.ident,
                                               sobj$cell_type)))
colnames(df_proportion) = c("orig.ident", "cell_type", "freq")

quantif = table(sobj$orig.ident) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("orig.ident", "nb_cells"))

aquarius::plot_barplot(df = df_proportion,
                       x = "orig.ident",
                       y = "freq",
                       fill = "cell_type",
                       position = ggplot2::position_fill()) +
  ggplot2::scale_fill_manual(name = "Cell type",
                             values = color_markers[levels(df_proportion$cell_type)],
                             breaks = levels(df_proportion$cell_type)) +
  ggplot2::geom_label(data = quantif, inherit.aes = FALSE,
                      aes(x = orig.ident, y = 1.05, label = nb_cells),
                      label.size = 0)
```


### Cell cycle

We annotate cells for cell cycle phase :

```{r cell_cycle2}
cc_columns = aquarius::add_cell_cycle(sobj = sobj,
                                      assay = "RNA",
                                      species_rdx = "mm",
                                      BPPARAM = cl)@meta.data[, c("Seurat.Phase", "Phase")]

sobj$Seurat.Phase = cc_columns$Seurat.Phase
sobj$cyclone.Phase = cc_columns$Phase
```



### Clustering

We make a highly resolutive clustering :

```{r clustering}
sobj = Seurat::FindNeighbors(sobj, reduction = "RNA_pca", dims = c(1:ndims))
sobj = Seurat::FindClusters(sobj, resolution = 1)

table(sobj$seurat_clusters)
```


# Visualization

## Cell type

We can visualize the cell type :

```{r see_cell_type, fig.width = 12, fig.height = 6, class.source = "fold-hide"}
tsne = Seurat::DimPlot(sobj, group.by = "cell_type",
                       reduction = paste0("RNA_pca_", ndims, "_tsne"), cols = color_markers) +
  Seurat::NoAxes() + ggplot2::ggtitle("tSNE") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

umap = Seurat::DimPlot(sobj, group.by = "cell_type",
                       reduction = paste0("RNA_pca_", ndims, "_umap"), cols = color_markers) +
  Seurat::NoAxes() + ggplot2::ggtitle("UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

tsne | umap
```

## Cell cycle

We can visualize the cell cycle, from Seurat :

```{r see_cc_Seurat, fig.width = 12, fig.height = 6, class.source = "fold-hide"}
tsne = Seurat::DimPlot(sobj, group.by = "Seurat.Phase",
                       reduction = paste0("RNA_pca_", ndims, "_tsne")) +
  Seurat::NoAxes() + ggplot2::ggtitle("tSNE") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

umap = Seurat::DimPlot(sobj, group.by = "Seurat.Phase",
                       reduction = paste0("RNA_pca_", ndims, "_umap")) +
  Seurat::NoAxes() + ggplot2::ggtitle("UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

tsne | umap
```


We can visualize the cell cycle, from cyclone :

```{r see_cc_cyclone, fig.width = 12, fig.height = 6, class.source = "fold-hide"}
tsne = Seurat::DimPlot(sobj, group.by = "cyclone.Phase",
                       reduction = paste0("RNA_pca_", ndims, "_tsne")) +
  Seurat::NoAxes() + ggplot2::ggtitle("tSNE") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

umap = Seurat::DimPlot(sobj, group.by = "cyclone.Phase",
                       reduction = paste0("RNA_pca_", ndims, "_umap")) +
  Seurat::NoAxes() + ggplot2::ggtitle("UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

tsne | umap
```


## Clusters

We visualize the clustering :

```{r see_clusters, fig.width = 12, fig.height = 6, class.source = "fold-hide"}
tsne = Seurat::DimPlot(sobj, group.by = "seurat_clusters", label = TRUE,
                       reduction = paste0("RNA_pca_", ndims, "_tsne")) +
  Seurat::NoAxes() + ggplot2::ggtitle("tSNE") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

umap = Seurat::DimPlot(sobj, group.by = "seurat_clusters", label = TRUE,
                       reduction = paste0("RNA_pca_", ndims, "_umap")) +
  Seurat::NoAxes() + ggplot2::ggtitle("UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

tsne | umap
```


## Gene expression

We visualize all cell types markers on the tSNE :

```{r plot_genes, fig.width = 12, fig.height = 50}
# markers = c(cell_markers)
# markers_vec = unlist(markers)
# names(markers_vec) = lapply(c(1:length(markers)), FUN = function(id) {
#   rep(names(markers)[id], length(markers[[id]]))
# }) %>% unlist()
# markers_vec = markers_vec[!duplicated(markers_vec)] # unique removes names
# markers_vec = markers_vec[markers_vec %in% rownames(sobj)]
# 
# plot_list = lapply(c(1:length(markers_vec)),
#                    FUN = function(gene_id) {
#                      one_gene = markers_vec[gene_id]
#                      one_pop = names(markers_vec[gene_id])
#                      p = Seurat::FeaturePlot(sobj, features = one_gene,
#                                              reduction = paste0("RNA_pca_", ndims, "_umap")) +
#                        ggplot2::labs(title = one_gene,
#                                      subtitle = one_pop) +
#                        ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
#                        ggplot2::theme(aspect.ratio = 1,
#                                       plot.subtitle = element_text(hjust = 0.5)) +
#                        Seurat::NoAxes()
#                      return(p)
#                    })
# 
# patchwork::wrap_plots(plot_list, ncol = 5)
```
## Cluster smoothing

```{r clustering_smoothing}
cluster_type = table(sobj$cell_type, sobj$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type = setNames(nm = names(cluster_type),
                        levels(sobj$cell_type)[cluster_type])

sobj$cluster_type = cluster_type[sobj$seurat_clusters]
```

## New clusters

We visualize the clustering :

```{r see_new.clusters, fig.width = 12, fig.height = 6, class.source = "fold-hide"}
tsne = Seurat::DimPlot(sobj, group.by = "cluster_type", label = TRUE,
                       reduction = paste0("RNA_pca_", ndims, "_tsne")) +
  Seurat::NoAxes() + ggplot2::ggtitle("tSNE") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

umap = Seurat::DimPlot(sobj, group.by = "cluster_type", label = TRUE,
                       reduction = paste0("RNA_pca_", ndims, "_umap")) +
  Seurat::NoAxes() + ggplot2::ggtitle("UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

tsne | umap
```
## Cluster markers

# ```{r markers_clusters, eval=FALSE, echo=FALSE}
# cluster.markers <- Seurat::FindMarkers(sobj, ident.1 = c(85), ident.2 = c(136,86,96), min.pct = 0.25)
# head(cluster.markers, n = 30)
# ```

# Save

We save the annotated and filtered Seurat object :

```{r save_sobj_filtered_annotated}
saveRDS(sobj, file = paste0(out_dir, "/datasets/", sample_name, "_sobj_filtered.rds"))
```


# R session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```
