---
title: "Comparison 3h vs 18h dissociation"
subtitle: "SCs_analysis"
author: "Pernelle"
date: '2024-03-25'
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

```{r clean, echo = FALSE}
rm(list = ls())
```

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <= Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
```

# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "SC"
out_dir = "."
```

We load the dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/../6_SCs_extraction/", save_name, "_sobj.rds"))
sobj
```

and sample information :

```{r load_sample_info}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "dissociation_methods_metadata_sample_info.rds"))
sample_info
```

This is the projection of interest :

```{r name2D}
name2D = "harmony_50_umap"
```

We load metadata :

```{r load_metadata}
color_markers = readRDS(paste0(out_dir, "/../1_metadata/", "dissociation_methods_metadata_color_markers.rds"))
cell_markers = readRDS(paste0(out_dir, "/../1_metadata/", "dissociation_methods_metadata_cell_markers.rds"))
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "dissociation_methods_metadata_sample_info.rds"))
```

We visualize clusters :

```{r cluster_vizualization, fig.width = 12, fig.height = 8, eval = FALSE}
Seurat::DimPlot(sobj, reduction = name2D, group.by ="cell_type", label = FALSE, cols = color_markers, pt.size = 0.5) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 0.8)
```

We change the resolution of the dataset :

```{r resolution}
sobj = Seurat::FindClusters(sobj, resolution = 1.5)

Seurat::DimPlot(sobj, reduction = name2D, group.by ="seurat_clusters", label = TRUE, pt.size = 0.5, label.size = 15) +
  Seurat::NoAxes() +
  Seurat::NoLegend() +
  ggplot2::theme(aspect.ratio = 0.8)
```

# Cell types

We define markers for each population of interest :

```{r cell_markers, class.source = "fold-hide"}
cell_markers_SCs = list(
  "Myelinating schwann cells" = c("Sox10", "Plp1", "S100b", "Lgi4", "Mpz","Pmp22","Mal","Egr2"),
  "Non myelinating schwann cells" = c("Sox10", "Plp1", "S100b", "Lgi4","Gap43","Ngfr","Scn7a"),
  "NA" = c("Fabp4", "Ebf1", "Mast4", "Rgs5","Epcam"))

lengths(cell_markers_SCs)
```

Cell populations annotation with the new resolution : 

Here are custom colors for each cell type :

```{r color_markers_SC, fig.width = 8, fig.height = 0.8, class.source = "fold-hide"}
color_markers_SCs = aquarius::c30_palette[c(1:length(cell_markers))]
names(color_markers_SCs) = names(cell_markers_SCs)

data.frame(cell_type = names(cell_markers_SCs),
           color = unlist(color_markers_SCs)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank(),
                 axis.text.x = element_text(angle = 10))
```

We annotate cells for cell type, with the new normalized expression matrix :

```{r cell_type_2, class.source = "fold-hide", time_it = FALSE}
score_columns = grep(x = colnames(sobj@meta.data), pattern = "^score", value = TRUE)
sobj@meta.data[, score_columns] = NULL
sobj$cell_type = NULL

sobj = aquarius::cell_annot_custom(sobj,
                                   newname = "cell_type",
                                   markers = cell_markers_SCs,
                                   use_negative = TRUE,
                                   add_score = TRUE,
                                   verbose = TRUE)

sobj$cell_type = factor(sobj$cell_type, levels = names(cell_markers_SCs))

colnames(sobj@meta.data) = stringr::str_replace_all(string = colnames(sobj@meta.data),
                                                    pattern = " ",
                                                    replacement = "_")

table(sobj$cell_type)
```

## Cluster smoothing

```{r clustering_smoothing}
cluster_type = table(sobj$cell_type, sobj$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type = setNames(nm = names(cluster_type),
                        levels(sobj$cell_type)[cluster_type])

sobj$cluster_type = cluster_type[sobj$seurat_clusters]
```

We visualize the new clusters :

```{r vizualisation, fig.width = 12, fig.height = 8}
Seurat::Idents(sobj) = sobj$cluster_type
Seurat::DimPlot(subset(sobj,idents = "NA", invert = TRUE), reduction = name2D, group.by ="cluster_type", label = FALSE, pt.size = 0.5, cols = color_markers_SCs) +
  Seurat::NoLegend() +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 0.8)

```

and split by origin :

```{r plot_split_dimred, fig.width = 12, fig.height = 10}
sobj$condition_diss = paste0(sobj$condition,"_",sobj$dissociation)


sobj$condition_diss = factor(sobj$condition_diss, levels = c("HLS_3hLib","cNFs_3hLib","HLS_18hLib","cNFs_18hLib"))

plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "condition_diss",
                                        group_by = "cluster_type",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = color_markers_SCs,
                                        bg_color = "gray85", bg_pt_size = 0.3, main_pt_size = 0.3)

patchwork::wrap_plots(plot_list, ncol = 4) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right") &
  guides(color = guide_legend(override.aes = list(size = 3), title = "Cluster"))
```

```{r barplot_count_SCs_subtypes, fig.width = 5, fig.height = 6}
Seurat::Idents(sobj) = sobj$cluster_type
sobj = subset(sobj, idents = "NA", invert = TRUE)
sobj$dissociation_condition = paste0(sobj$dissociation, "_",sobj$condition)

sobj$dissociation_condition = factor(sobj$dissociation_condition, levels = c("3hLib_HLS","3hLib_cNFs","18hLib_HLS","18hLib_cNFs"))

quantif = table(sobj$dissociation_condition) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Sample", "nb_cells"))

plot_list = table(sobj$dissociation_condition, sobj$cluster_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c( "Sample", "Cell Type", "Number")) %>%
  aquarius::plot_barplot(.,
                         x = "Sample", y = "Number", fill = "Cell Type",
                         position = position_fill()) +
  ggplot2::scale_fill_manual(values = unlist(color_markers_SCs),
                             breaks = names(color_markers_SCs),
                             name = "Cell Type") +
  ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.5),
                              labels = paste0(seq(0, 100, by = 50)," %"),
                              expand = ggplot2::expansion(add = c(0, 0.05))) +
  ggplot2::geom_label(data = quantif, inherit.aes = FALSE,
                      aes(x = .data$Sample, y = 1.05, label = .data$nb_cells),
                      label.size = 0, size = 8) +
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.line.x = element_line(colour = "lightgray"),
                 text = element_text(size = 15),
                 legend.position = "none")
plot_list
```

```{r collagen_score, fig.width = 11.77, fig.height = 10}
Col_genes = c("Col1a1", "Col1a2","Col2a1","Col3a1","Col4a1","Col4a2","Col4a3", "Col4a4","Col4a5","Col4a6","Col5a1","Col5a2","Col5a3","Col6a1","Col6a2","Col6a3", "Col6a4","Col6a5","Col6a6","Col7a1","Col8a1", "Col8a2","Col9a1", "Col9a2","Col9a3","Col10a1","Col11a1", "Col11a2","Col12a1","Col13a1","Col14a1","Col15a1","Col16a1","Col17a1","Col18a1","Col19a1","Col20a1","Col22a1","Col23a1","Col24a1","Col25a1","Col26a1","Col27a1","Col28a1")

sobj$Col_score = Seurat::AddModuleScore(sobj, features = list(Col_genes))$Cluster1

colors = c("chartreuse2","firebrick2","chartreuse4","firebrick4")

sobj$condition_diss = factor(sobj$condition_diss, levels = c("HLS_3hLib","cNFs_3hLib","HLS_18hLib","cNFs_18hLib"))

Seurat::VlnPlot(sobj, features = c("Col_score"), group.by ="sample_type", split.by = "cluster_type", pt.size = 0, ncol =1) + ggplot2::geom_boxplot(position = position_dodge(0.90), width=0.15, outlier.shape = NA, fatten = NULL) + ggplot2::stat_summary(fun=mean, geom="point", color="black", position = position_dodge(0.90), width=0.3)

plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        color_by = "tdTomato",
                                        split_by = "condition_diss",
                                        bg_pt_size = 0.5,
                                        main_pt_size = 0.5)

plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot = one_plot +
    ggplot2::scale_color_gradientn(colors = aquarius::color_gene,
                                 name = "Expression level") +
  # ggplot2::theme(legend.key.height = unit(5,"mm"),
  #                legend.key.width = unit(5,"mm"),
  #                legend.title = element_text(size = 20),
  #                legend.text = element_text(size = 20)) +
  ggplot2::theme(aspect.ratio = 0.8)
  return(one_plot)
})

names(plot_list) = levels(sobj$condition_diss)
plot_list = plot_list[c("HLS_3hLib","cNFs_3hLib","HLS_18hLib","cNFs_18hLib")]
patchwork::wrap_plots(plot_list, ncol = 4)
```

```{r diff_expression, fig.width = 20, fig.height = 7.5}
sobj$interest = paste0(sobj$condition_diss,"_",sobj$cluster_type)

Seurat::Idents(sobj) = sobj$interest

mark_3h = Seurat::FindMarkers(sobj, ident.1 = "cNFs_3hLib_Non myelinating schwann cells", ident.2="HLS_3hLib_Non myelinating schwann cells") %>%
  as.data.frame() %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::filter(avg_logFC>0)

mark_18h = Seurat::FindMarkers(sobj, ident.1 = "cNFs_18hLib_Non myelinating schwann cells", ident.2="HLS_18hLib_Non myelinating schwann cells") %>%
  as.data.frame() %>%
  dplyr::arrange(avg_logFC) %>%
  dplyr::filter(avg_logFC>0)

diff_3h = setdiff(rownames(mark_3h), rownames(mark_18h))
diff_18h = setdiff(rownames(mark_18h), rownames(mark_3h))

ggvenn::ggvenn(list("3h" = rownames(mark_3h),
                    "18h" = rownames(mark_18h)),
               text_size = 10, show_percentage = FALSE) +
  ggplot2::labs(title = "Comparison ECM genes between SG and other nmSCs") +
  ggplot2::theme(plot.title = element_text(hjust = 0.5, size = 20))
```

# Save

We save the Seurat object :

```{r save_sobj}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_sobj.rds"))
saveRDS(sample_info, file = paste0(out_dir, "/", save_name, "_sample_info.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```







