---
title: "Axonal guidance cues genes"
subtitle: "Analysis heatmaps + violin plots"
author: "Pernelle"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

```{r clean, echo = FALSE}
rm(list = ls())
```

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <= Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(org.Mm.eg.db)
library(ComplexHeatmap)
```


# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "SC"
out_dir = "."
```


We load the dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/../5_annotation/", save_name, "_sobj.rds"))
sobj
```

and sample information :

```{r load_sample_info}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "Tom_metadata_sample_info.rds"))
sample_info
```

This is the projection of interest :

```{r name2D}
name2D = "harmony_46_umap"
```

We load metadata :

```{r load_metadata}
color_markers = readRDS(paste0(out_dir, "/../1_metadata/", "Tom_metadata_color_markers.rds"))
cell_markers = readRDS(paste0(out_dir, "/../1_metadata/", "Tom_metadata_cell_markers.rds"))
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "Tom_metadata_sample_info.rds"))
```

We visualize clusters :

```{r clustering, fig.width = 12, fig.height = 8, eval = FALSE}
cell_markers_SCs_fin = list("Other nmSCs" = c("Sox10", "Plp1", "S100b", "Lgi4","Gap43","Ngfr","Scn7a"),
                            #"Subepidermal glia" = c("Aqp1","Plac9","Igfbp7","Ldlrad4","Apod","Ndnf","Ifi47","Gja1","Pcdh20","Fez1","Lyz2"),
                            "Subepidermal glia" = c("Gja1","Aqp1","Igfbp7","Fez1","Plac9"),
                            "Lanceolate glia" = c("Aqp1", "Homer2", "Npy","Cav1","Robo2","Cavin2"),
                            "Proliferating nmSCs" = c("Birc5","Top2a","Mki67","H1f1"),
                            "mSCs" = c("Mbp","Mpz","Pmp22"))
color_markers_SCs_fin = c("red4","darkorchid","olivedrab1","chocolate3","dodgerblue3")
names(color_markers_SCs_fin) = names(cell_markers_SCs_fin)

Seurat::DimPlot(sobj, reduction = name2D, group.by ="cluster_type_SCs", label = FALSE, cols = color_markers_SCs_fin, pt.size = 0.5) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 0.8)
```


```{r annot_global, class.source = "fold-hide"}
cell_markers_SCs = list(
  "Myelinating schwann cells" = c("Sox10", "Plp1", "S100b", "Lgi4", "Mpz","Pmp22","Mal","Egr2"),
  "Non myelinating schwann cells" = c("Sox10", "Plp1", "S100b", "Lgi4","Gap43","Ngfr","Scn7a"),
  "NA" = c("Fabp4", "Ebf1", "Mast4", "Rgs5","Epcam"))

color_markers_SCs = aquarius::c30_palette[c(1:length(cell_markers))]
names(color_markers_SCs) = names(cell_markers_SCs)

Seurat::DimPlot(sobj, reduction = name2D, group.by ="cluster_type", label = FALSE, cols = color_markers, pt.size = 0.5) +
  Seurat::NoLegend() +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 0.8)
```


```{r axonal_guidance_cues_list, fig.width =50, fig.height = 7}
all_genes = sobj@assays[["RNA"]]@counts %>% rownames()
Semaphorins = grep(all_genes, pattern = "^Sema", value = TRUE)
Plexins = grep(all_genes, pattern = "^Plxn", value = TRUE)
Netrins = grep(all_genes, pattern = "^Ntn", value = TRUE)
Neurotrophic = c("Ngf","Gdnf","Ntf3","Ntf4","Ntf5")

EFN = grep(all_genes, pattern = "^Efn", value = TRUE)
EPH = grep(all_genes, pattern = "^Eph", value = TRUE)
Plexins = c("Plxna1","Plxna2","Plxna3","Plxna4a","Plxnb1","Plxnb2","Plxnb3","Plxnc1","Plxnd1")
Slit = c("Slit1","Slit2","Slit3")

Axon_genes = c(Semaphorins, Plexins, Netrins, Neurotrophic, EFN, EPH, Plexins, Slit)
sobj$Axon_score = Seurat::AddModuleScore(sobj, features = list(Axon_genes))$Cluster1
```

```{r collagen_score_vln_plot, fig.width = 22, fig.height = 7}
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365"))

colors = c("purple2","orchid2","firebrick2","orange2", "chartreuse2","cadetblue2")

Seurat::Idents(sobj) = sobj$cluster_type_SCs

Seurat::VlnPlot(subset(sobj, idents = "Proliferating nmSCs", invert = TRUE), features = c("Axon_score"), split.by= "sample_type", group.by ="cluster_type_SCs", cols = colors, pt.size = 0, ncol =1)  + ggplot2::geom_boxplot(position = position_dodge(0.90), width=0.15, outlier.shape = NA, fatten = NULL) + ggplot2::stat_summary(fun=mean, geom="point", color="black", position = position_dodge(0.90), width=0.3)
```

```{r collagen_score_feature_plot_split, fig.width = 22, fig.height = 7}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        color_by = "Axon_score",
                                        split_by = "sample_type",
                                        bg_pt_size = 0.5,
                                        main_pt_size = 0.5)

plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot = one_plot +
    ggplot2::scale_color_gradientn(colors = aquarius::color_gene,
                                 name = "Expression level") +
  # ggplot2::theme(legend.key.height = unit(5,"mm"),
  #                legend.key.width = unit(5,"mm"),
  #                legend.title = element_text(size = 20),
  #                legend.text = element_text(size = 20)) +
  ggplot2::theme(aspect.ratio = 0.8)
  return(one_plot)
})

names(plot_list) = levels(sobj$sample_type)
plot_list = plot_list[c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365")]
patchwork::wrap_plots(plot_list, ncol = 3)
```






```{r collagen_score_vln_plot, fig.width = 22, fig.height = 7}
Axon_genes = c(Semaphorins, Plexins, Netrins, Neurotrophic, EFN, EPH, Plexins, Slit)
sobj$GOBP_NEGATIVE_REGULATION_OF_AXONOGENESIS = Seurat::AddModuleScore(sobj, features = list(GOBP_NEGATIVE_REGULATION_OF_AXONOGENESIS))$Cluster1


sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365"))

colors = c("purple2","orchid2","firebrick2","orange2", "chartreuse2","cadetblue2")

Seurat::Idents(sobj) = sobj$cluster_type_SCs

Seurat::VlnPlot(subset(sobj, idents = "Proliferating nmSCs", invert = TRUE), features = c("GOBP_NEGATIVE_REGULATION_OF_AXONOGENESIS"), split.by= "sample_type", group.by ="cluster_type_SCs", cols = colors, pt.size = 0, ncol =1)  + ggplot2::geom_boxplot(position = position_dodge(0.90), width=0.15, outlier.shape = NA, fatten = NULL) + ggplot2::stat_summary(fun=mean, geom="point", color="black", position = position_dodge(0.90), width=0.3)
```

```{r diff_expression, fig.width =50, fig.height = 7}
sobj$condition_SCs = paste0(sobj$cluster_type_SCs, "_",sobj$sample_type)
Seurat::Idents(sobj) = sobj$condition_SCs

markP90 = Seurat::FindMarkers(sobj, ident.1 = c("Subepidermal glia_flfl_cNFs_P90"), ident.2 = c("Subepidermal glia_flfl_HLS_P90")) %>%
  as.data.frame() %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0)
```


```{r axonal_cues}
library(readxl)
axonal_cues = read_excel("/home/aurelien/Bureau/scRNAseq_Tom_cells/6_analyse/Axonal_guidance_cues.xlsx")

SemaP90 = intersect(axonal_cues$Semaphorins, rownames(markP90))
Sema = unique(c(SemaP90))

NetrinP90 = intersect(axonal_cues$'Netrin-related', rownames(markP90))
Netrin = unique(c(NetrinP90))

RoboP90 = intersect(axonal_cues$'Robos', rownames(markP90))
Robo = unique(c(RoboP90))

SlitP90 = intersect(axonal_cues$'Slits', rownames(markP90))
Slit = unique(c(SlitP90))

EfnP90 = intersect(axonal_cues$'Ephrins', rownames(markP90))
Efn = unique(c(EfnP90))

EphP90 = intersect(axonal_cues$'Ephrin_receptors', rownames(markP90))
Eph = unique(c(EphP90))

NFP90 = intersect(axonal_cues$'Neurotrophic_factors', rownames(markP90))
NF = unique(c(NFP90))

CAMP90 = intersect(axonal_cues$'CAMs', rownames(markP90))
CAM = unique(c(CAMP90))

PlxP90 = intersect(axonal_cues$'Plexins', rownames(markP90))
Plx = unique(c(PlxP90))

NrplP90 = intersect(axonal_cues$'Neuropilins', rownames(markP90))
Nrpl = unique(c(NrplP90))

```

```{r dotplot_axonal_cues_sum_up, fig.width = 20, fig.height = 7.5}
Seurat::Idents(sobj) = sobj$sample_type
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365"))
# subsobj1 = subset(sobj, idents = c("ctrl_HS_P90","ctrl_HS_P365"), invert = TRUE)
# subsobj1$sample_type = factor(subsobj1$sample_type, levels = c("flfl_HLS_P90","flfl_cNFs_P90","flfl_cNFs_P365","flfl_HLS_P365"))

Seurat::Idents(sobj) = sobj$cluster_type_SCs
Seurat::DotPlot(subset(sobj,idents = c("Subepidermal glia")),
                features = c(Sema, Netrin, Robo, Slit, Efn, Eph, NF, CAM, Plx, Nrpl),
                group.by = "sample_type", dot.scale = 15) +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  ggplot2::theme(axis.title = element_blank(),
                 legend.key.height = unit(10,"mm"),
                 legend.key.width = unit(10,"mm"),
                 legend.title = element_text(size = 25),
                 legend.text = element_text(size = 25),
                 axis.text = element_text(size = 30),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggplot2::theme(aspect.ratio = 0.3)
```


```{r diff_expression_downregulated, fig.width =50, fig.height = 7}
sobj$condition_SCs = paste0(sobj$cluster_type_SCs, "_",sobj$sample_type)
Seurat::Idents(sobj) = sobj$condition_SCs

markP365 = Seurat::FindMarkers(sobj, ident.1 = c("Subepidermal glia_flfl_cNFs_P365"), ident.2 = c("Subepidermal glia_flfl_HLS_P365")) %>%
  as.data.frame() %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0) 

markP90 = Seurat::FindMarkers(sobj, ident.1 = c("Subepidermal glia_flfl_cNFs_P90"), ident.2 = c("Subepidermal glia_flfl_HLS_P90")) %>%
  as.data.frame() %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC < 0)
```


```{r axonal_cues_downregulated}
library(readxl)
axonal_cues = read_excel("/home/aurelien/Bureau/scRNAseq_Tom_cells/6_analyse/Axonal_guidance_cues.xlsx")

SemaP90 = intersect(axonal_cues$Semaphorins, rownames(markP90))
SemaP365 = intersect(axonal_cues$Semaphorins, rownames(markP365))
Sema = unique(c(SemaP90,SemaP365))

NetrinP90 = intersect(axonal_cues$'Netrin-related', rownames(markP90))
NetrinP365 = intersect(axonal_cues$'Netrin-related', rownames(markP365))
Netrin = unique(c(NetrinP90,NetrinP365))

RoboP90 = intersect(axonal_cues$'Robos', rownames(markP90))
RoboP365 = intersect(axonal_cues$'Robos', rownames(markP365))
Robo = unique(c(RoboP90,RoboP365))

SlitP90 = intersect(axonal_cues$'Slits', rownames(markP90))
SlitP365 = intersect(axonal_cues$'Slits', rownames(markP365))
Slit = unique(c(SlitP90,SlitP365))

EfnP90 = intersect(axonal_cues$'Ephrins', rownames(markP90))
EfnP365 = intersect(axonal_cues$'Ephrins', rownames(markP365))
Efn = unique(c(EfnP90,EfnP365))

EphP90 = intersect(axonal_cues$'Ephrin_receptors', rownames(markP90))
EphP365 = intersect(axonal_cues$'Ephrin_receptors', rownames(markP365))
Eph = unique(c(EphP90,EphP365))

NFP90 = intersect(axonal_cues$'Neurotrophic_factors', rownames(markP90))
NFP365 = intersect(axonal_cues$'Neurotrophic_factors', rownames(markP365))
NF = unique(c(NFP90,NFP365))

CAMP90 = intersect(axonal_cues$'CAMs', rownames(markP90))
CAMP365 = intersect(axonal_cues$'CAMs', rownames(markP365))
CAM = unique(c(CAMP90,CAMP365))

PlxP90 = intersect(axonal_cues$'Plexins', rownames(markP90))
PlxP365 = intersect(axonal_cues$'Plexins', rownames(markP365))
Plx = unique(c(PlxP90,PlxP365))

NrplP90 = intersect(axonal_cues$'Neuropilins', rownames(markP90))
NrplP365 = intersect(axonal_cues$'Neuropilins', rownames(markP365))
Nrpl = unique(c(NrplP90,NrplP365))

```

```{r dotplot_down_regulated_axonal_cues_sum_up, fig.width = 20, fig.height = 7.5}
Seurat::Idents(sobj) = sobj$sample_type
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365"))
subsobj1 = subset(sobj, idents = c("ctrl_HS_P90","ctrl_HS_P365"), invert = TRUE)
subsobj1$sample_type = factor(subsobj1$sample_type, levels = c("flfl_HLS_P90","flfl_cNFs_P90","flfl_cNFs_P365","flfl_HLS_P365"))

Seurat::Idents(subsobj1) = subsobj1$cluster_type_SCs
Seurat::DotPlot(subset(subsobj1,idents = c("Subepidermal glia")),
                features = c(Sema, Netrin, Robo, Slit, Efn, Eph, NF, CAM, Plx, Nrpl),
                group.by = "sample_type", dot.scale = 15) +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  ggplot2::theme(axis.title = element_blank(),
                 legend.key.height = unit(10,"mm"),
                 legend.key.width = unit(10,"mm"),
                 legend.title = element_text(size = 25),
                 legend.text = element_text(size = 25),
                 axis.text = element_text(size = 30),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggplot2::theme(aspect.ratio = 0.3)
```



```{r diff_expression_other_nmSCs, fig.width =50, fig.height = 7}
sobj$condition_SCs = paste0(sobj$cluster_type_SCs, "_",sobj$sample_type)
Seurat::Idents(sobj) = sobj$condition_SCs

markP365 = Seurat::FindMarkers(sobj, ident.1 = c("Other nmSCs_flfl_cNFs_P365"), ident.2 = c("Other nmSCs_flfl_HLS_P365")) %>%
  as.data.frame() %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) 

markP90 = Seurat::FindMarkers(sobj, ident.1 = c("Other nmSCs_flfl_cNFs_P90"), ident.2 = c("Other nmSCs_flfl_HLS_P90")) %>%
  as.data.frame() %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0)
```


```{r axonal_cues_other_nmSCs}
SemaP90 = intersect(axonal_cues$Semaphorins, rownames(markP90))
SemaP365 = intersect(axonal_cues$Semaphorins, rownames(markP365))
Sema = unique(c(SemaP90,SemaP365))

NetrinP90 = intersect(axonal_cues$'Netrin-related', rownames(markP90))
NetrinP365 = intersect(axonal_cues$'Netrin-related', rownames(markP365))
Netrin = unique(c(NetrinP90,NetrinP365))

RoboP90 = intersect(axonal_cues$'Robos', rownames(markP90))
RoboP365 = intersect(axonal_cues$'Robos', rownames(markP365))
Robo = unique(c(RoboP90,RoboP365))

SlitP90 = intersect(axonal_cues$'Slits', rownames(markP90))
SlitP365 = intersect(axonal_cues$'Slits', rownames(markP365))
Slit = unique(c(SlitP90,SlitP365))

EfnP90 = intersect(axonal_cues$'Ephrins', rownames(markP90))
EfnP365 = intersect(axonal_cues$'Ephrins', rownames(markP365))
Efn = unique(c(EfnP90,EfnP365))

EphP90 = intersect(axonal_cues$'Ephrin_receptors', rownames(markP90))
EphP365 = intersect(axonal_cues$'Ephrin_receptors', rownames(markP365))
Eph = unique(c(EphP90,EphP365))

NFP90 = intersect(axonal_cues$'Neurotrophic_factors', rownames(markP90))
NFP365 = intersect(axonal_cues$'Neurotrophic_factors', rownames(markP365))
NF = unique(c(NFP90,NFP365))

CAMP90 = intersect(axonal_cues$'CAMs', rownames(markP90))
CAMP365 = intersect(axonal_cues$'CAMs', rownames(markP365))
CAM = unique(c(CAMP90,CAMP365))

PlxP90 = intersect(axonal_cues$'Plexins', rownames(markP90))
PlxP365 = intersect(axonal_cues$'Plexins', rownames(markP365))
Plx = unique(c(PlxP90,PlxP365))

NrplP90 = intersect(axonal_cues$'Neuropilins', rownames(markP90))
NrplP365 = intersect(axonal_cues$'Neuropilins', rownames(markP365))
Nrpl = unique(c(NrplP90,NrplP365))

```

```{r dotplot_axonal_cues_other_nmSCs_sum_up, fig.width = 20, fig.height = 7.5}
Seurat::Idents(sobj) = sobj$sample_type
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365"))
subsobj1 = subset(sobj, idents = c("ctrl_HS_P90","ctrl_HS_P365"), invert = TRUE)
subsobj1$sample_type = factor(subsobj1$sample_type, levels = c("flfl_HLS_P90","flfl_cNFs_P90","flfl_cNFs_P365","flfl_HLS_P365"))

Seurat::Idents(subsobj1) = subsobj1$cluster_type_SCs
Seurat::DotPlot(subset(subsobj1,idents = c("Other nmSCs")),
                features = c(Sema, Netrin, Robo, Slit, Efn, Eph, NF, CAM, Plx, Nrpl),
                group.by = "sample_type", dot.scale = 15) +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  ggplot2::theme(axis.title = element_blank(),
                 legend.key.height = unit(10,"mm"),
                 legend.key.width = unit(10,"mm"),
                 legend.title = element_text(size = 25),
                 legend.text = element_text(size = 25),
                 axis.text = element_text(size = 30),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggplot2::theme(aspect.ratio = 0.3)
```

```{r genes_exploration_feature_plot_split, fig.width = 22, fig.height = 7}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        color_by = "Nes",
                                        split_by = "sample_type",
                                        bg_pt_size = 0.5,
                                        main_pt_size = 0.5)

plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot = one_plot +
    ggplot2::scale_color_gradientn(colors = aquarius::color_gene,
                                 name = "Expression level") +
  # ggplot2::theme(legend.key.height = unit(5,"mm"),
  #                legend.key.width = unit(5,"mm"),
  #                legend.title = element_text(size = 20),
  #                legend.text = element_text(size = 20)) +
  ggplot2::theme(aspect.ratio = 0.8)
  return(one_plot)
})

names(plot_list) = levels(sobj$sample_type)
plot_list = plot_list[c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365")]
patchwork::wrap_plots(plot_list, ncol = 3)
```


```{r diff_expression, fig.width =50, fig.height = 7}
sobj$condition_SCs = paste0(sobj$cluster_type, "_",sobj$sample_type)
Seurat::Idents(sobj) = sobj$condition_SCs

markP90 = Seurat::FindMarkers(sobj, ident.1 = c("Non myelinating schwann cells_flfl_cNFs_P90"), ident.2 = c("Non myelinating schwann cells_flfl_HLS_P90")) %>%
  as.data.frame() %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0)
```


```{r axonal_cues}
library(readxl)
axonal_cues = read_excel("/home/aurelien/Bureau/scRNAseq_Tom_cells/6_analyse/Axonal_guidance_cues.xlsx")

SemaP90 = intersect(axonal_cues$Semaphorins, rownames(markP90))
Sema = unique(c(SemaP90))

NetrinP90 = intersect(axonal_cues$'Netrin-related', rownames(markP90))
Netrin = unique(c(NetrinP90))

RoboP90 = intersect(axonal_cues$'Robos', rownames(markP90))
Robo = unique(c(RoboP90))

SlitP90 = intersect(axonal_cues$'Slits', rownames(markP90))
Slit = unique(c(SlitP90))

EfnP90 = intersect(axonal_cues$'Ephrins', rownames(markP90))
Efn = unique(c(EfnP90))

EphP90 = intersect(axonal_cues$'Ephrin_receptors', rownames(markP90))
Eph = unique(c(EphP90))

NFP90 = intersect(axonal_cues$'Neurotrophic_factors', rownames(markP90))
NF = unique(c(NFP90))

CAMP90 = intersect(axonal_cues$'CAMs', rownames(markP90))
CAM = unique(c(CAMP90))

PlxP90 = intersect(axonal_cues$'Plexins', rownames(markP90))
Plx = unique(c(PlxP90))

NrplP90 = intersect(axonal_cues$'Neuropilins', rownames(markP90))
Nrpl = unique(c(NrplP90))
```

```{r dotplot_axonal_cues_sum_up, fig.width = 20, fig.height = 7.5}
Seurat::Idents(sobj) = sobj$sample_type
sobj = subset(sobj, idents = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90"))
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90"))

Seurat::Idents(sobj) = sobj$sample_type
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90"))
# subsobj1 = subset(sobj, idents = c("ctrl_HS_P90","ctrl_HS_P365"), invert = TRUE)
# subsobj1$sample_type = factor(subsobj1$sample_type, levels = c("flfl_HLS_P90","flfl_cNFs_P90","flfl_cNFs_P365","flfl_HLS_P365"))

Seurat::Idents(sobj) = sobj$cluster_type
Seurat::DotPlot(subset(sobj,idents = c("Non myelinating schwann cells")),
                features = c(Sema, Netrin, Robo, Slit, Efn, Eph, NF, CAM, Plx, Nrpl),
                group.by = "sample_type", dot.scale = 15) +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  ggplot2::theme(axis.title = element_blank(),
                 legend.key.height = unit(10,"mm"),
                 legend.key.width = unit(10,"mm"),
                 legend.title = element_text(size = 25),
                 legend.text = element_text(size = 25),
                 axis.text = element_text(size = 30),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggplot2::theme(aspect.ratio = 0.3)
```


```{r split_object_SG}
sobj$condition_SCs = paste0(sobj$cluster_type, "_",sobj$sample_type)
Seurat::Idents(sobj) = sobj$condition_SCs

mark_nmSCs_P365 = Seurat::FindMarkers(sobj, ident.1 = c("Non myelinating schwann cells_flfl_cNFs_P365"), ident.2 = c("Non myelinating schwann cells_flfl_HLS_P365")) %>%
  as.data.frame() %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) 

mark_nmSCs_P90 = Seurat::FindMarkers(sobj, ident.1 = c("Non myelinating schwann cells_flfl_cNFs_P90"), ident.2 = c("Non myelinating schwann cells_flfl_HLS_P90")) %>%
  as.data.frame() %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(avg_logFC > 0) 

unlist(rownames(mark_nmSCs_P365))
```


We keep all up-regulated or down-regulated genes.+

```{r DeSEQ2_results}
up_genes = rownames(mark_nmSCs_P90)

```



## ORA and GSEA

We make over-representation analysis for each group of genes. We load gene sets from MSigDB :

```{r gene_sets}
gene_sets = aquarius::get_gene_sets()
# gene_sets_full = gene_sets$gene_sets_full
# long_names = gene_sets$long_names
gene_sets = gene_sets$gene_sets
head(gene_sets)
```

We get gene name and gene ID correspondence :

```{r gene_corresp}
gene_corresp = read.csv("/home/aurelien/Bureau/scRNAseq_Tom_cells/gene_corresp.csv", row.names = 1)
head(gene_corresp)
```

## Up-regulated genes

We explore enrichment in gene sets :

```{r enrichr_results_up, fig.width = 12, fig.height = 8}

genes_of_interest = up_genes

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp[, c("NAME", "ENSEMBL")],
                                        gene_sets = gene_sets[, c("gs_name","ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = "Up-regulated genes in cNFs cells compared to HLS - P90")

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

We run a GSEA with differentially expressed gene sets :

```{r gsea_P90_cNFs_vs_HLS, fig.width = 15, fig.height = 90}

sobj$condition_SCs = paste0(sobj$cell_type, "_",sobj$sample_type)
Seurat::Idents(sobj) = sobj$condition_SCs

ranked_gene_list = aquarius::run_foldchange(
  Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
  group1 = names(which(sobj$`condition_SCs` == c("Non myelinating schwann cells_flfl_HLS_P90"))),
  group2 = names(which(sobj$`condition_SCs` == c("Non myelinating schwann cells_flfl_cNFs_P90"))))

names(ranked_gene_list) = gene_corresp$ENSEMBL

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                  GSEA_p_val_thresh = 1)

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

We visualize the GSEA using cluster profiler:

```{r cluster_profiler, fig.width = 12, fig.height = 6}
AXON = unlist(grep(gsea_results@result[["ID"]], pattern = "AXON", value = TRUE))
enrichplot::gseaplot2(gsea_results, geneSetID=AXON)
AXON = AXON %>%
  as.data.frame()

AXON = AXON[c(1,2,3,4,8,9,11,13,16,17,22,29,30,33,34,35,36,41,46,48,50),]

AXON = droplevels(AXON)

gsea_results@geneSets[["GOBP_POSITIVE_REGULATION_OF_AXON_EXTENSION"]]

enrichplot::gseaplot2(gsea_results, geneSetID="KEGG_AXON_GUIDANCE") #BIEN
enrichplot::gseaplot2(gsea_results, geneSetID="REACTOME_REGULATION_OF_COMMISSURAL_AXON_PATHFINDING_BY_SLIT_AND_ROBO") #nul
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_AXONOGENESIS_INVOLVED_IN_INNERVATION") #nul
enrichplot::gseaplot2(gsea_results, geneSetID="GOCC_AXONAL_GROWTH_CONE") #nul
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_NEGATIVE_REGULATION_OF_AXON_EXTENSION_INVOLVED_IN_AXON_GUIDANCE") #moyen
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_NEGATIVE_REGULATION_OF_AXON_EXTENSION_INVOLVED_IN_AXON_GUIDANCE")
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_REGULATION_OF_AXON_GUIDANCE")
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_AXON_DEVELOPMENT") # BIEN
enrichplot::gseaplot2(gsea_results, geneSetID="REACTOME_SEMA3A_PAK_DEPENDENT_AXON_REPULSION")
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_AXON_EXTENSION") # pas trop mal
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_REGULATION_OF_AXON_DIAMETER") #nul
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_NEGATIVE_REGULATION_OF_AXON_EXTENSION") #bien mais contradictoire
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_PERIPHERAL_NERVOUS_SYSTEM_AXON_REGENERATION") # nul
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_REGULATION_OF_AXONOGENESIS") #pas trop mal
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_POSITIVE_REGULATION_OF_AXONOGENESIS") #pas trop mal
enrichplot::gseaplot2(gsea_results, geneSetID="GOMF_AXON_GUIDANCE_RECEPTOR_ACTIVITY")#nul
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_POSITIVE_REGULATION_OF_AXON_EXTENSION") #bof moyen
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_NEGATIVE_REGULATION_OF_AXONOGENESIS") #bien mais contradictoire
enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_AXONAL_FASCICULATION") # nul
enrichplot::gseaplot2(gsea_results, geneSetID="REACTOME_P75NTR_REGULATES_AXONOGENESIS") #nul
enrichplot::gseaplot2(gsea_results, geneSetID="KEGG_AXON_GUIDANCE")


enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_POSITIVE_REGULATION_OF_AXON_EXTENSION")

enrichplot::gseaplot2(gsea_results, geneSetID="GOBP_REGULATION_OF_NEURON_PROJECTION_DEVELOPMENT")
```

```{r cluster_profiler, fig.width = 12, fig.height = 6}
GOBP_AXON_EXTENSION = gsea_results@geneSets[["GOBP_AXON_EXTENSION"]]
GOBP_AXON_EXTENSION = as.vector(gene_corresp[gene_corresp$ENSEMBL %in% GOBP_AXON_EXTENSION, "NAME"])
GOBP_AXON_EXTENSION = intersect(rownames(mark_nmSCs_P90), GOBP_AXON_EXTENSION)


GOBP_POSITIVE_REGULATION_OF_AXONOGENESIS = gsea_results@geneSets[["GOBP_POSITIVE_REGULATION_OF_AXONOGENESIS"]]
GOBP_POSITIVE_REGULATION_OF_AXONOGENESIS = as.vector(gene_corresp[gene_corresp$ENSEMBL %in% GOBP_POSITIVE_REGULATION_OF_AXONOGENESIS, "NAME"])
GOBP_POSITIVE_REGULATION_OF_AXONOGENESIS = intersect(rownames(mark_nmSCs_P90), GOBP_POSITIVE_REGULATION_OF_AXONOGENESIS)


KEGG_AXON_GUIDANCE = gsea_results@geneSets[["KEGG_AXON_GUIDANCE"]]
KEGG_AXON_GUIDANCE = as.vector(gene_corresp[gene_corresp$ENSEMBL %in% KEGG_AXON_GUIDANCE, "NAME"])
KEGG_AXON_GUIDANCE = intersect(rownames(mark_nmSCs_P90), KEGG_AXON_GUIDANCE)

KEGG_AXON_GUIDANCE = gsea_results@geneSets[["KEGG_AXON_GUIDANCE"]]
KEGG_AXON_GUIDANCE = as.vector(gene_corresp[gene_corresp$ENSEMBL %in% list, "NAME"])
intersect(rownames(mark_nmSCs_P90), list2)

GOBP_AXON_DEVELOPMENT = gsea_results@geneSets[["GOBP_AXON_DEVELOPMENT"]]
GOBP_AXON_DEVELOPMENT = as.vector(gene_corresp[gene_corresp$ENSEMBL %in% GOBP_AXON_DEVELOPMENT, "NAME"])
GOBP_AXON_DEVELOPMENT = intersect(rownames(mark_nmSCs_P90), GOBP_AXON_DEVELOPMENT)

GOBP_NEGATIVE_REGULATION_OF_AXONOGENESIS = gsea_results@geneSets[["GOBP_AXON_DEVELOPMENT"]]
GOBP_NEGATIVE_REGULATION_OF_AXONOGENESIS = as.vector(gene_corresp[gene_corresp$ENSEMBL %in% GOBP_NEGATIVE_REGULATION_OF_AXONOGENESIS, "NAME"])
GOBP_NEGATIVE_REGULATION_OF_AXONOGENESIS = intersect(rownames(mark_nmSCs_P90), GOBP_NEGATIVE_REGULATION_OF_AXONOGENESIS)
```

We run a GSEA with complete gene sets :

```{r gsea_test_clusters_2_vs_1, fig.width = 12, fig.height = 60}
sobj$condition_SCs = paste0(sobj$cell_type, "_",sobj$sample_type)
Seurat::Idents(sobj) = sobj$condition_SCs

ranked_gene_list = aquarius::run_foldchange(
  Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
  group1 = names(which(sobj$`condition_SCs` == c("Non myelinating schwann cells_flfl_HLS_P90"))),
  group2 = names(which(sobj$`condition_SCs` == c("Non myelinating schwann cells_flfl_cNFs_P90"))))

names(ranked_gene_list) = ifelse(is.na(gene_corresp[names(ranked_gene_list), "ENTREZID"]),
                                 yes = names(ranked_gene_list),
                                 no = gene_corresp[names(ranked_gene_list), "ENTREZID"])

gene_set_all= unique(gene_sets$gs_name)
gene_sets_df = gene_sets[gene_sets$gs_name %in%
                           gene_set_all, c("gs_name", "entrez_gene")]
gene_sets_df = na.omit(gene_sets_df)
gene_sets_df = gene_sets_df[gene_sets_df$gs_name %in% names(which(table(gene_sets_df$gs_name) > 5)), ]

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gene_sets_df,
                                  GSEA_p_val_thresh = 1)

gsea_results_p <- gsea_results@result %>%
  dplyr::filter(pvalue < 0.05)

list_results[["All"]]$gsea_up = gsea_results@result %>%
  dplyr::filter(pvalue < 0.05, enrichmentScore > 0)

list_results[["All"]]$gsea_down = gsea_results@result %>%
  dplyr::filter(pvalue < 0.05, enrichmentScore < 0)


gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))

```


```{r heatmap_set_sobj_SG, fig.width = 22, fig.height = 6}
Seurat::Idents(sobj) = sobj$cell_type
subsobj_HM = subset(sobj, idents = "Non myelinating schwann cells")

subsobj_HM$sample_type = factor(subsobj_HM$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","flfl_cNFs_P365","flfl_HLS_P365","ctrl_HS_P365"))
```


```{r heatmap_set_genes_SG, fig.width = 22, fig.height = 6}
genes_heatmap = list(group01 = KEGG_AXON_GUIDANCE)


features_group = sapply(c(1:length(genes_heatmap)),
                        FUN = function(i) {
                          rep(x = names(genes_heatmap)[i],
                              length = length(genes_heatmap[[i]]))}) %>% unlist() %>% as.vector()
features = unique(unlist(genes_heatmap))

```

```{r heatmap_set_matrix_SG, fig.width = 22, fig.height = 6}

mat_expression = Seurat::GetAssayData(subsobj_HM, assay = "RNA", slot = "data")
mat_expression = mat_expression[features, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
dim(mat_expression)
```

```{r heatmap_set_annotation_SG, fig.width = 22, fig.height = 6}
list_colors = list()
list_colors[["expression"]] = aquarius::color_cnv

# change colors !!
list_colors[["sample_type"]] = c(ctrl_HS_P90 = "white",
                                 flfl_HLS_P90 = "white",
                                 flfl_cNFs_P90 = "white",
                                 flfl_cNFs_P365 = "white",
                                 flfl_HLS_P365 = "white",
                                 ctrl_HS_P365 = "white")
list_colors[["sample_identifier"]] = setNames(nm = sample_info$sample_identifier,
                                              sample_info$color)
list_colors[["cell_cycle_phase"]] = aquarius::gg_color_hue(3)
names(list_colors[["cell_cycle_phase"]]) = c("G0 / G1", "G2M", "S")
list_colors[["percent_mito"]] = circlize::colorRamp2(breaks = seq(from = min(subsobj_HM$percent.mt),
                                                                  to = max(sobj$percent.mt),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
list_colors[["nFeature_RNA"]] = circlize::colorRamp2(breaks = seq(from = min(subsobj_HM$nFeature_RNA),
                                                                  to = max(subsobj_HM$nFeature_RNA),
                                                                  length.out = 9),
                                                     colors = RColorBrewer::brewer.pal(name = "Greys", n = 9))
```

```{r heatmap_set_color_SG, fig.width = 22, fig.height = 6}
name_side = "right"
ha_top = HeatmapAnnotation(
  group = anno_block(gp = gpar(fill = list_colors[["sample_type"]]),
                     labels = sort(unique(subsobj_HM$sample_type))
                     ),
  sample_identifier = subsobj_HM$sample_identifier,
  col = list(sample_identifier = list_colors[["sample_identifier"]]),
  annotation_name_side = name_side)

ha_bottom = HeatmapAnnotation(
  percent_mito = subsobj_HM$percent.mt,
  nFeature_RNA = subsobj_HM$nFeature_RNA,
  # cell_cycle_phase = as.factor(subsobj_HM$Seurat.Phase) %>% `levels<-`(c("G0 / G1", "G2M", "S")),
  col = list(percent_mito = list_colors[["percent_mito"]],
             nFeature_RNA = list_colors[["nFeature_RNA"]]),
             # cell_cycle_phase = list_colors[["cell_cycle_phase"]]),
  annotation_name_side = name_side)
```

```{r heatmap_set_SG, fig.width = 22, fig.height = 6}
ht = ComplexHeatmap::Heatmap(
  as.matrix(mat_expression),
  # Expression colors
  col = list_colors[["expression"]],
  heatmap_legend_param = list(title = "Expression", at = c(0, 1), 
                              labels = c("low", "high")),
  # Annotation
  top_annotation = ha_top,
  bottom_annotation = ha_bottom,
  # Cells grouping
  column_split = subsobj_HM$sample_type,
  column_gap = unit(2, "mm"),
  column_title = NULL,
  # Cells order
  column_order = order(subsobj_HM$sample_type),
  cluster_columns = FALSE, #faire un clustering des cellules
  show_column_dend = FALSE, #montrer l'arbre du clustering (dendrogramme)
  # Cells label
  show_column_names = FALSE,
  # Gene grouping
  row_split = features_group,
  row_gap = unit(2, "mm"),
  row_title = NULL,
  cluster_rows = FALSE, #clusteriser les gènes à peu près co-exprimés
  show_row_dend = FALSE, #montrer l'arbre du clustering (dendrogramme)
  # Gene labels
  row_labels = features,
  show_row_names = TRUE,
  row_names_side = name_side,
  row_names_gp = grid::gpar(fontsize = 13), #customiser la taille du texte des noms des lignes
  # Visual aspect
  show_heatmap_legend = TRUE,
  border = TRUE,
  use_raster = FALSE)
```

```{r draw_heatmap_SG, fig.width = 22, fig.height = 15}
ComplexHeatmap::draw(ht, # legend_grouping = "original",
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```