---
title: "Annotation_SCs_subtypes"
author: "Pernelle"
date: '2024-03-14'
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

```{r clean, echo = FALSE}
rm(list = ls())
```

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <= Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
```

# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "SC"
out_dir = "."
```

We load the dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/../4_zoomSC/", save_name, "_sobj.rds"))
sobj
```

and sample information :

```{r load_sample_info}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "Tom_metadata_sample_info.rds"))
sample_info
```

This is the projection of interest :

```{r name2D}
name2D = "harmony_46_umap"
```

We load metadata :

```{r load_metadata}
color_markers = readRDS(paste0(out_dir, "/../1_metadata/", "Tom_metadata_color_markers.rds"))
cell_markers = readRDS(paste0(out_dir, "/../1_metadata/", "Tom_metadata_cell_markers.rds"))
sample_info = readRDS(paste0(out_dir, "/../1_metadata/", "Tom_metadata_sample_info.rds"))
```

We visualize clusters :

```{r cluster_vizualization, fig.width = 12, fig.height = 8, eval = FALSE}
Seurat::DimPlot(sobj, reduction = name2D, group.by ="cluster_type", label = FALSE, cols = color_markers, pt.size = 0.5) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 0.8)
```

We change the resolution of the dataset :

```{r resolution}
sobj = Seurat::FindClusters(sobj, resolution = 1.5)

Seurat::DimPlot(sobj, reduction = name2D, group.by ="seurat_clusters", label = TRUE, pt.size = 0.5, label.size = 15) +
  Seurat::NoAxes() +
  Seurat::NoLegend() +
  ggplot2::theme(aspect.ratio = 0.8)
```

# Cell types

We define markers for each population of interest :

```{r cell_markers, class.source = "fold-hide"}
cell_markers_SCs = list(
  "Myelinating schwann cells" = c("Sox10", "Plp1", "S100b", "Lgi4", "Mpz","Pmp22","Mal","Egr2"),
  "Non myelinating schwann cells" = c("Sox10", "Plp1", "S100b", "Lgi4","Gap43","Ngfr","Scn7a"),
  "NA" = c("Fabp4", "Ebf1", "Mast4", "Rgs5","Epcam"))

lengths(cell_markers_SCs)
```

Cell populations annotation with the new resolution : 

Here are custom colors for each cell type :

```{r color_markers_SC, fig.width = 8, fig.height = 0.8, class.source = "fold-hide"}
color_markers_SCs = aquarius::c30_palette[c(1:length(cell_markers))]
names(color_markers_SCs) = names(cell_markers_SCs)

data.frame(cell_type = names(color_markers_SCs),
           color = unlist(color_markers_SCs)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank(),
                 axis.text.x = element_text(angle = 10))
```

We annotate cells for cell type, with the new normalized expression matrix :

```{r cell_type_2, class.source = "fold-hide", time_it = FALSE}
score_columns = grep(x = colnames(sobj@meta.data), pattern = "^score", value = TRUE)
sobj@meta.data[, score_columns] = NULL
sobj$cell_type = NULL

sobj = aquarius::cell_annot_custom(sobj,
                                   newname = "cell_type",
                                   markers = cell_markers_SCs,
                                   use_negative = TRUE,
                                   add_score = TRUE,
                                   verbose = TRUE)

sobj$cell_type = factor(sobj$cell_type, levels = names(cell_markers_SCs))

colnames(sobj@meta.data) = stringr::str_replace_all(string = colnames(sobj@meta.data),
                                                    pattern = " ",
                                                    replacement = "_")

table(sobj$cell_type)
```

## Cluster smoothing

```{r clustering_smoothing}
cluster_type = table(sobj$cell_type, sobj$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type = setNames(nm = names(cluster_type),
                        levels(sobj$cell_type)[cluster_type])

sobj$cluster_type = cluster_type[sobj$seurat_clusters]
```

We visualize the new clusters :

```{r clustering, fig.width = 12, fig.height = 8, eval = FALSE}
Seurat::DimPlot(sobj, reduction = name2D, group.by ="cluster_type", label = FALSE, cols = color_markers, pt.size = 0.5) +
  Seurat::NoLegend() +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 0.8)
```

and split by origin :

```{r plot_split_dimred, fig.width = 12, fig.height = 10}
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365"))

plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "sample_type",
                                        group_by = "cluster_type",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = color_markers,
                                        bg_color = "gray85", bg_pt_size = 0.3, main_pt_size = 0.3)

patchwork::wrap_plots(plot_list, ncol = 3) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right") &
  guides(color = guide_legend(override.aes = list(size = 3), title = "Cluster"))
```

```{r barplot_count, fig.width = 11.77, fig.height = 10}
Seurat::Idents(sobj) = sobj$cluster_type

quantif = table(sobj$sample_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Sample", "nb_cells"))

plot_list = table(sobj$sample_type, sobj$cluster_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c( "Sample", "Cell Type", "Number")) %>%
  aquarius::plot_barplot(.,
                         x = "Sample", y = "Number", fill = "Cell Type",
                         position = position_fill()) +
  ggplot2::scale_fill_manual(values = unlist(color_markers),
                             breaks = names(color_markers),
                             name = "Cell Type") +
  ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.5),
                              labels = paste0(seq(0, 100, by = 50)," %"),
                              expand = ggplot2::expansion(add = c(0, 0.05))) +
  ggplot2::geom_label(data = quantif, inherit.aes = FALSE,
                      aes(x = .data$Sample, y = 1.05, label = .data$nb_cells),
                      label.size = 0, size = 8) +
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.line.x = element_line(colour = "lightgray"),
                 text = element_text(size = 15),
                 legend.position = "none")
plot_list
```

```{r dotplot_SCs_markers, fig.width = 20, fig.height = 7.5}

Seurat::Idents(sobj) = sobj$cluster_type

Seurat::DotPlot(sobj,
                features = c( "Ngfr","Gap43","Scn7a","Pmp22","Mbp","Mpz","Sox10","S100b","Plp1"),
                group.by = "cluster_type", dot.scale = 15) +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  ggplot2::theme(axis.title = element_blank(),
                 legend.key.height = unit(10,"mm"),
                 legend.key.width = unit(10,"mm"),
                 legend.title = element_text(size = 25),
                 legend.text = element_text(size = 25),
                 axis.text = element_text(size = 30),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggplot2::theme(aspect.ratio = 0.5)
```

```{r feature_plot_Tomato, fig.width = 12, fig.height = 7}

Tomato = Seurat::FeaturePlot(sobj, features = c("tdTomato"), reduction = name2D, label = FALSE, pt.size = 0.5) +
  Seurat::NoAxes() +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene,
                                 name = "Expression level") +
  ggplot2::theme(legend.key.height = unit(10,"mm"),
                 legend.key.width = unit(10,"mm"),
                 legend.title = element_text(size = 20),
                 legend.text = element_text(size = 20)) +
  ggplot2::theme(aspect.ratio = 0.8) 

ARN_mt = Seurat::FeaturePlot(sobj, features = c("percent.mt"), reduction = name2D, label = FALSE, pt.size = 0.5) +
  Seurat::NoAxes() +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene,
                                 name = "Expression level") +
  ggplot2::theme(legend.key.height = unit(10,"mm"),
                 legend.key.width = unit(10,"mm"),
                 legend.title = element_text(size = 20),
                 legend.text = element_text(size = 20)) +
  ggplot2::theme(aspect.ratio = 0.8) 

Tomato | ARN_mt
```

# SC subtypes

```{r cell_markers_version_fine, class.source = "fold-hide"}
cell_markers_SCs_fin = list("Other nmSCs" = c("Sox10", "Plp1", "S100b", "Lgi4","Gap43","Ngfr","Scn7a"),
                            #"Subepidermal glia" = c("Aqp1","Plac9","Igfbp7","Ldlrad4","Apod","Ndnf","Ifi47","Gja1","Pcdh20","Fez1","Lyz2"),
                            "Subepidermal glia" = c("Gja1","Aqp1","Igfbp7","Fez1","Plac9"),
                            "Lanceolate glia" = c("Aqp1", "Homer2", "Npy","Cav1","Robo2","Cavin2"),
                            "Proliferating nmSCs" = c("Birc5","Top2a","Mki67","H1f1"),
                            "mSCs" = c("Mbp","Mpz","Pmp22"))

lengths(cell_markers_SCs_fin)
```


```{r color_markers_SC_subtypes, fig.width = 8, fig.height = 0.8, class.source = "fold-hide"}
color_markers_SCs_fin = c("red4","darkorchid","olivedrab1","chocolate3","dodgerblue3")
names(color_markers_SCs_fin) = names(cell_markers_SCs_fin)

data.frame(cell_type = names(color_markers_SCs_fin),
           color = unlist(color_markers_SCs_fin)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers_SCs_fin), breaks = names(color_markers_SCs_fin)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank(),
                 axis.text.x = element_text(angle = 10))
```


```{r new_annotation, class.source = "fold-hide"}
sobj = aquarius::cell_annot_custom(sobj,
                                   newname = "SC_subtypes",
                                   markers = cell_markers_SCs_fin,
                                   use_negative = TRUE,
                                   add_score = TRUE,
                                   verbose = TRUE)

colnames(sobj@meta.data) = stringr::str_replace_all(string = colnames(sobj@meta.data),
                                                    pattern = " ",
                                                    replacement = "_")

sobj$SC_subtypes = factor(sobj$SC_subtypes, levels = names(cell_markers_SCs_fin))

table(sobj$SC_subtypes) 


Seurat::DimPlot(sobj, reduction = name2D, group.by ="SC_subtypes", label = TRUE, pt.size = 0.5, label.size = 15) +
  Seurat::NoAxes() +
  Seurat::NoLegend() +
  ggplot2::theme(aspect.ratio = 0.8)

```


```{r cluster_smoothing_SC_subtypes, fig.width = 12, fig.height = 8}
cluster_type_SCs = table(sobj$SC_subtypes, sobj$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cluster_type_SCs = setNames(nm = names(cluster_type_SCs),
                            levels(sobj$SC_subtypes)[cluster_type_SCs])

sobj$cluster_type_SCs = cluster_type_SCs[sobj$seurat_clusters]


Seurat::DimPlot(sobj, reduction = name2D, group.by ="cluster_type_SCs", label = FALSE, pt.size = 0.5, cols = color_markers_SCs_fin) +
  Seurat::NoLegend() +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 0.8)

```


```{r barplot_count_beautiful_prep}
sobj$cluster_type_SCs = factor(sobj$cluster_type_SCs, levels = c("Lanceolate glia","mSCs","Proliferating nmSCs","Subepidermal glia","Other nmSCs"))

quantif = table(sobj$sample_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Sample", "nb_cells"))

quantif2 = table(sobj$cluster_type_SCs, sobj$sample_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Cell_type", "Sample_type","nb_cell")) %>%
  dplyr::mutate(Cell_type = as.character(Cell_type)) %>%
  dplyr::group_by(Sample_type) %>%
  dplyr::mutate(prop_cell = nb_cell/sum(nb_cell)) %>%
  dplyr::mutate(prop_cum =  cumsum(nb_cell/sum(nb_cell))) %>%
  dplyr::mutate(prop_cum_mid = prop_cum) %>%
  as.data.frame()

quantif3 = lapply(unique(quantif2$Sample_type), FUN = function(Sample_type) {
  subquantif2 = quantif2[quantif2$Sample_type == Sample_type, ]
  
  for (i in c(1:nrow(subquantif2))) {
    if (i == 1) {
      subquantif2[i, "prop_cum_mid"] = subquantif2[i, "prop_cum"]/2
    } else {
      subquantif2[i, "prop_cum_mid"] = (subquantif2[i - 1, "prop_cum"] + subquantif2[i, "prop_cum"])/2
    }
  }
  
  return(subquantif2)
}) %>% do.call(rbind.data.frame, .) %>%
  dplyr::mutate(prop_cell = 100*round(prop_cell, 4))

quantif3 = quantif3[-c(3,18,23,28),]
```

```{r barplot_count_SCs_subtypes, fig.width = 11.77, fig.height = 10}
sobj$cluster_type_SCs = factor(sobj$cluster_type_SCs, levels = c("Lanceolate glia","mSCs","Proliferating nmSCs","Subepidermal glia","Other nmSCs"))

table(sobj$sample_type, sobj$cluster_type_SCs) %>%
  as.data.frame.table() %>%
  `colnames<-`(c( "Sample", "Cell Type", "Number")) %>%
  aquarius::plot_barplot(.,
                         x = "Sample", y = "Number", fill = "Cell Type",
                         position = position_fill()) +
  # Fill color
  ggplot2::scale_fill_manual(values = unlist(color_markers_SCs_fin),
                             breaks = names(color_markers_SCs_fin),
                             name = "Cell Type") +
  # Y axis style
  ggplot2::scale_y_continuous(breaks = seq(0, 1, by = 0.5),
                              labels = paste0(seq(0, 100, by = 50)," %"),
                              expand = ggplot2::expansion(add = c(0, 0.05))) +
  # Number of cells by sample
  ggplot2::geom_label(data = quantif, inherit.aes = FALSE,
                      aes(x = Sample, y = 1.05, label = nb_cells),
                      label.size = 0, size = 5) +
  # Proportion of cell types by sample
  ggplot2::geom_label(data = quantif3,
                      inherit.aes = FALSE, fill = "white",
                      label.size = 0, size = 5,
                      aes(x = Sample_type,
                          y = 1 - prop_cum_mid,
                          label = paste0(prop_cell, "%"))) +
  # Global style
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.line.x = element_line(colour = "lightgray"),
                 text = element_text(size = 15),
                 legend.position = "none")
  
```


```{r dotplot_SC_subtypes_markers, fig.width = 25, fig.height = 7.5}
Seurat::Idents(sobj) = sobj$cluster_type_SCs
sobj$cluster_type_SCs = factor(sobj$cluster_type_SCs, levels = c("mSCs","Proliferating nmSCs","Lanceolate glia","Subepidermal glia","Other nmSCs"))

Seurat::DotPlot(sobj,
                features = c("Mpz","Mbp","Pmp22","Mki67","Top2a","Birc5","Cavin2","Cav1","Homer2","Npy","Nes","Ndnf","Apod","Gja1","Aqp1","Igfbp7","Fez1","Lyz2","Plac9","Gap43","Ngfr","Scn7a"),
                group.by = "cluster_type_SCs", dot.scale = 15) +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  ggplot2::theme(axis.title = element_blank(),
                 legend.key.height = unit(10,"mm"),
                 legend.key.width = unit(10,"mm"),
                 legend.title = element_text(size = 25),
                 legend.text = element_text(size = 25),
                 axis.text = element_text(size = 30),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggplot2::theme(aspect.ratio = 0.4)
```


```{r plot_split_dimred_SCs_subtypes, fig.width = 14, fig.height = 8}
sobj$sample_type = factor(sobj$sample_type, levels = c("ctrl_HS_P90","flfl_HLS_P90","flfl_cNFs_P90","ctrl_HS_P365","flfl_HLS_P365","flfl_cNFs_P365"))

plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "sample_type",
                                        group_by = "cluster_type_SCs",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = color_markers_SCs_fin,
                                        bg_color = "gray85", bg_pt_size = 0.3, main_pt_size = 0.1)

patchwork::wrap_plots(plot_list, ncol = 3) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right") &
  guides(color = guide_legend(override.aes = list(size = 3), title = "Cluster")) &
  ggplot2::theme(aspect.ratio = 0.8)
```


# Save

We save the Seurat object :

```{r save_sobj}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_sobj.rds"))
saveRDS(sample_info, file = paste0(out_dir, "/", save_name, "_sample_info.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```







